---
title: "intro"
output:
  bookdown::html_document2:
    code_folding: hide

---

```{r, include = FALSE}
source(here::here("start_rmd.R"))
```
# Load data

```{r}
#withr::with_dir(here::here(), {
tar_load(timeseries)
#}
#)
```


# Get some information 

```{r}
summary(timeseries)
```

- I want histograms et summary of quantitative variables 

```{r}
mask_num_var <- map_lgl(timeseries, is.numeric)

lg_timeseries_num <- timeseries[, mask_num_var] %>%
  pivot_longer(cols = everything(),
    names_to = "variable",
    values_to = "values")
```

```{r hist_num}
#| hist_num, echo = FALSE, fig.width = 10,
#| fig.cap = "This is a long long
#|   long long caption."

lg_timeseries_num %>%
  ggplot(aes(x = values)) +
  geom_histogram() +
  facet_wrap(vars(variable), scales = "free")
```

```{r}
map2_dfr(
  timeseries[, mask_num_var], c(4, 2, 3, 3, 3, 8, 3, 4, 2, 2),
  ~summary_distribution(.x, na.rm = TRUE) %>% formatC(., digits = .y),
  .id = "variable") %>%
kable(., caption = "Descriptive statistics of Global timeseries", label = "d-num")
```

# Site characteristics

- Site are described by siteid  
- Gathering info such as protocol, year range, location (gps, region),

```{r}
var_site <- c("species", "date", "year", "month", "quarter", "protocol",
  "protocol_detail", "sampledlength_m", "speciesstatus", "unitabundance",
  "unitbiomass", "country", "abundance", "biomass", "ecoregion", "region",
  "latitude", "longitude", "province", "waterbody")

var_localisation <- get_var_localisation()

var_protocol <- get_var_protocol()
```

```{r}
x <- rep(list(c(runif(1e3), rep(NA, 1000))), 100)
bench::mark(
  map(x, get_unique_values),
  map(x, get_unique_values_c)
)[c("expression", "min", "median", "itr/sec", "n_gc")]
```


```{r}
tar_load(site_desc_loc)
```


- Check that location are unique for all site 

```{r}
map(site_desc_loc, ~ tibble(
    no_unique = .x[.x %in% c("no_unique")] %>% length,
    na = .x[.x %in% NA] %>% length)) %>%
  enframe(., name = "variable", value = "count") %>%
  unnest(cols = c(count)) %>%
  kable(., caption = "Nb sites which have NAs or no unique values")
```

- Some sites have no or several provinces, region and waterbody, looks
  manageable

```{r}
site_desc_loc %>%
  filter(waterbody == "no_unique")

timeseries %>%
  filter(siteid == filter(site_desc_loc, waterbody == "no_unique")$siteid[1:6]) %>%
  group_by(siteid) %>%
  summarise(unique(waterbody))
```

- Well, there are some problems with the waterbody name but I do not think that it is
  really important


- I will described:
  - Quantitative: 
    - [x] Total species number, total, median, mean, sd, min, max 
    - [x] year: min, max, median, range, sd, cover,
    - [x] month: min, max, median, mean, sd 
    - [x] sampledlength_m:  min, max, median, mean, sd
  - Qualitative:
    - [x] protocol: unique
    - [x] protocol_detail: unique
    - [x] speciesstatus: unique
    - [x] unitabundance: unique
    - [x] unitbiomass: unique

## Species number and species traits 

```{r}
tar_load(species_number_op)
tar_load(abun_rich_op)
tar_load(species_number_site)
tar_load(species_status)
```



## Fish operation

```{r}
tar_load(op_protocol)
```

```{r}
# Fish operation description
check_quali <- map_lgl(op_protocol[,!colnames(op_protocol) %in% "date"], ~all(.x != "no_unique", na.rm = TRUE))
```

- Make sure that each operation is described by a unique protocol, sampling
  effort, measurement unit, date: `r all(check_quali)`
  - This is not the case for `r names(check_quali)[!check_quali]`, let's check 

```{r}
op_protocol %>%
  filter(unitbiomass == "no_unique")

timeseries %>%
  filter(op_id == filter(op_protocol, unitbiomass ==
      "no_unique")$op_id[2]) %>%
select(species, abundance, biomass, unitbiomass)

timeseries %>%
  filter(is.na(unitbiomass), !is.na(biomass))


# Check if there are biomass unit without biomass measurement:
all((!is.na(timeseries$unitbiomass) & is.na(timeseries$biomass)) == FALSE)
# Yes, but not important
```

- So, when biomass unit has not been reported, it means that biomass has not
  been recorded. For site and operation description, we keep the unique non-na
  values 

```{r, cache = FALSE}
tar_load(site_protocol_quali)
tar_load(site_protocol_quanti)
```


```{r}
unique(site_protocol_quali$protocol)
unique(site_protocol_quali$protocol_detail)

op_protocol %>%
  filter(siteid == filter(site_protocol_quali, protocol ==
      "no_unique")$siteid[14])
```
- Ok, there are 
  `r site_protocol_quali[site_protocol_quali$protocol == "no_unique",] %>% nrow`
  sites that do not have an unique protocol in a site

- There are 
  `r site_protocol_quali[site_protocol_quali$unitabundance == "no_unique",] %>% nrow`
  sites that do not have an unique unit of abundance measurement in a site


## Measurements 

- I save a data with only `op_id`, `species`, `abundance` and `biomass`

```{r}
tar_load(measurement)
tar_load(op_protocol)
tar_load(site_protocol_quanti)
```

# Map 

```{r}
site_protocol_loc <- site_desc_loc %>%
  select(siteid, latitude, longitude, ecoregion) %>%
  left_join(species_number_site %>% select(siteid, species_tot_nb), by = "siteid") %>%
  left_join(site_protocol_quali %>% select(siteid, protocol, protocol_detail, unitabundance), by = "siteid")
```

```{r}
site_protocol_loc_sf <- st_as_sf(site_protocol_loc, coords = c("longitude", "latitude"),
  crs = 4326)
```

- See ecoregion:
  (idea: do a panel with the 6 ecoregion to better see species number)

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = site_protocol_loc_sf, aes(size =
        species_tot_nb), shape = 1) +
    ggtitle("Total number of species by site")
```

- We have plenty of site that have a constant protocol and unit of abundance. 

```{r}
#https://www.graphpad.com/support/faq/graph-tip-dont-use-a-log-scale-on-a-bar-graph/
site_protocol_loc %>%
  ggplot(aes(x = protocol, fill = ecoregion)) +
  geom_bar() +
  labs(x = "Protocol", y = "Number of sites") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
site_protocol_loc %>%
  ggplot(aes(x = unitabundance, fill = ecoregion)) +
  geom_bar() +
  labs(x = "Abundance unit", y = "Number of sites") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

- Let's see them on a map

```{r}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = site_protocol_loc_sf, aes(color = protocol)) +
    xlab("Longitude") + ylab("Latitude") +
    ggtitle("Protocol")
```

```{r}
ggplot(data = world) +
    geom_sf() +
    geom_sf(data = site_protocol_loc_sf, aes(color = unitabundance)) +
    ggtitle("Unit of abundance")
```

```{r}
site_protocol_quanti %>%
  filter(variable == "year") %>%
  ggplot(aes(x = max -min )) +
  geom_histogram() +
  labs(x = "Year span (years)", y = "Number of sites") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
ti <- site_protocol_quanti %>%
  filter(variable == "year") %>%
  left_join(site_protocol_loc, by = "siteid") %>%
  mutate(
    span = max - min,
    cat_span = cut(span, c(0, 10, 20, 30, 40, 70), right = TRUE,
#      labels=c("<=10","10<x<=20","20<x<=30","30<x<=40", "40<x<=70"),
      include.lowest=TRUE
    ),
    cat_completness = cut(n / span, c(0, .1, .25, .5, .75, 1), right = TRUE,
#      labels=c("<=10","10<x<=20","20<x<=30","30<x<=40", "40<x<=70"),
      include.lowest=TRUE
    ),
  )
```
- Completeness

```{r}
ti %>%
  ggplot(aes(x = cat_completness, fill = ecoregion)) +
  geom_bar() +
  labs(x = "Completness (nb sampling years / year span)", y = "Number of sites")
```

```{r}
ti %>%
  ggplot(aes(x = cat_span, fill = ecoregion)) +
  geom_bar() +
  labs(x = "Year span", y = "Number of sites")
```

- Year baseline:

```{r}
site_protocol_quanti %>%
  filter(variable == "year") %>%
  ggplot(aes(x = min )) +
  geom_histogram() +
  labs(x = "First year of monitoring", y = "Number of sites") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# Abundance data

```{r}
abun_check <- measurement %>%
  group_by(op_id) %>%
  summarise(
    enframe(
      c(
        summary_distribution(abundance, na.rm = TRUE),
        sum = sum(abundance, na.rm = TRUE)
        )
      ),
    .groups = "drop"
  )
abun_check <- abun_check %>%
  ungroup() %>%
  left_join(op_protocol %>% select(op_id, unitabundance), by = "op_id") %>%
  pivot_wider(names_from = "name", values_from = "value")
```
```{r}
abun_check %>%
  ggplot(aes(x = sum)) +
  geom_histogram() +
  labs(y = "Count", x = "Total abundance", title = "Total abundance by sampling
    events") +
  facet_wrap(vars(unitabundance), scales = "free_x")
```

```{r}
total_abundance_distribution <- abun_check %>%
  group_by(unitabundance) %>%
  summarise(enframe(summary_distribution(sum)), .groups = "drop") %>%
  pivot_wider(names_from = "name", values_from = "value")

total_abundance_distribution %>%
  kable(., label = "tot-abun-dist", caption = "Distribution of total abundance
    by sampling event by unit of abundance")
```

# Trends

## Filter site

- TODO:
  - [] filter on month, quater 

- Keep the sites have been followed more than ten years

```{r}
mask_siteid_sampling <- site_protocol_quanti[
  site_protocol_quanti$variable == "year" &
    site_protocol_quanti$n >= 10,
  ]$siteid

mask_siteid_protocol <- site_protocol_quali[
  site_protocol_quali$unitabundance %in% c("Count", "Ind.100m2"), ]$siteid

mask_siteid <- mask_siteid_sampling[mask_siteid_sampling %in% mask_siteid_protocol]
```

```{r}
tar_load(op_protocol)
tar_load(abun_rich_op)

trends_data <- abun_rich_op %>%
  left_join(op_protocol, by = "op_id") %>%
  filter(siteid %in% mask_siteid) %>%
  mutate(
    log_total_abundance = log(total_abundance),
    log_species_nb = log(species_nb)
  )

rigal_test <- get_rigal_trajectory_classification(
  trends_data,
  y_var = "log_total_abundance",
  x_var = "year", site_id = "siteid")
```

```{r}
tar_load(site_desc_loc)
site_loc <- get_site_location_as_sf()

site_loc_trend <- site_loc %>%
  filter(siteid %in% rigal_test$siteid) %>%
  left_join(rigal_test %>%
    select(siteid, direction, shape_class, linear_slope,
      linear_slope_strd_error, linear_slope_pvalue),
  by = "siteid")

ggplot(data = world) +
    geom_sf() +
    geom_sf(data = site_loc_trend, aes(color =
        direction), shape = 1) +
    ggtitle("Total abundance trends")
hist(rigal_test$linear_slope, main = "Distribution of log total abundance
  temporal trend (slope)")
rigal_test %>%
  ggplot(aes(x = first_X_value, y = linear_slope)) + 
  geom_point() +
  geom_smooth(method = "auto") +
  labs(x = "Baseline year", "Temporal slope of total abundance (log)")
```

```{r}
rigal_test_richness <- get_rigal_trajectory_classification(
  trends_data,
  y_var = "log_species_nb",
  x_var = "year", site_id = "siteid")

site_loc_trend_richness <- site_loc %>%
  filter(siteid %in% rigal_test_richness$siteid) %>%
  left_join(rigal_test_richness %>%
    select(siteid, direction, shape_class, linear_slope,
      linear_slope_strd_error, linear_slope_pvalue),
  by = "siteid")

ggplot(data = world) +
    geom_sf() +
    geom_sf(data = site_loc_trend_richness, aes(color =
        direction), shape = 1) +
    ggtitle("Species richness temporal trends")

hist(rigal_test_richness$linear_slope,
  main = "Distribution of log species richness temporal trend (slope)")
hist(rigal_test_richness$first_X_value,
  main = "Distribution of first year of time sequence"
)
rigal_test_richness %>%
  ggplot(aes(x = first_X_value, y = linear_slope)) +
  geom_point() +
  geom_smooth(method = "auto") +
  labs(x = "Baseline year", "Temporal slope of species richness (log)")
```


```{r}
summary_loc_trends_data <- site_loc_trend_richness %>%
  group_by(ecoregion) %>%
  summarise(n = n())
site_loc_trend_richness %>%
  ggplot(aes(x = ecoregion, y = linear_slope)) +
  geom_violin() +
  geom_text(data = summary_loc_trends_data,
    aes(x = ecoregion, y = 0, label = paste0("n = ", n)),
    color = "red", fontface = 2, size = 5) +
  labs(y = "Temporal trends (slope) of species richness")
```

```{r}
trends_data %>%
  ggplot(aes(x = year, y = log_total_abundance)) +
  geom_line(aes(group = siteid)) +
  geom_smooth(aes(group = 1), size = 2, method = "lm", se = FALSE)

trends_data %>%
  ggplot(aes(x = year, y = log_species_nb)) +
  geom_line(aes(group = siteid)) +
  geom_smooth(aes(group = 1), size = 2, method = "lm", se = FALSE)
```

```{r, eval=FALSE}
library(brms)
fit1 <- brm(log_total_abundance ~ year + year^2,
            data = trends_data[trends_data$siteid == unique(trends_data$siteid)[1],], family = gaussian())

conditional_effects(fit1, surface = TRUE)
```

## Turnover

```{r, eval = FALSE}
tar_load(c(timeseries, site_protocol_quali))
selection_site <- op_protocol %>%
  filter(unitabundance %in% c("Count", "Ind.100m2")) %>%
  group_by(siteid) %>%
  summarise(n = n()) %>%
  filter(length(n) >= 10)
rm(test)

turnover <- timeseries %>%
  filter(siteid %in% selection_site$siteid) %>%
  nest_by(siteid) %>%
  mutate(
    turnover = list(try(turnover(
          df = data,
          time.var = "year",
          species.var = "species",
          abundance.var = "abundance",
          replicate.var = NA,
          metric = "total"
          ))),
  appearance = list(try(turnover(
        df = data,
        time.var = "year",
        species.var = "species",
        abundance.var = "abundance",
        replicate.var = NA,
        metric = "appearance"
        ))),
  disappearance = list(try(turnover(
        df = data,
        time.var = "year",
        species.var = "species",
        abundance.var = "abundance",
        replicate.var = NA,
        metric = "disappearance"
        )))
  )
  

turnover2 <- turnover[map_lgl(test$turnover, ~class(.x) != "try-error"),] %>%
  mutate(test = list(left_join(left_join(turnover, appearance, by = "year"),
        disappearance, by
        ="year"))
  )
#  unnest(cols = c(turnover, appearance, disappearance), names_repair = "universal")

turnover[map_lgl(turnover$turnover, ~class(.x) != "try-error"),] %>%
  ungroup() %>%
  mutate(test2 = map(list(turnover, appearance, disappearance), ~.x %>% reduce(left_join, by
        = "year")))

# Turnover to control for number of year between two time steps 
turnover3 <- map(c("turnover", "appearance", "disappearance"),
  ~turnover[map_lgl(turnover$turnover, ~class(.x) != "try-error"), c("siteid", .x)] %>%
    ungroup() %>%
    unnest(cols = .x)
) %>%
reduce(left_join, by = c("siteid", "year"))


```




# Population

```{r}
tar_load(toy_dataset)
```

```{r, eval=FALSE}
#devtools::install_github("Zoological-Society-of-London/rlpi", dependencies=TRUE)
library(rlpi)
CalcLPI(Species = toy_dataset$species, ID = toy_dataset$siteid, Year = toy_dataset$year,
  Popvalue = toy_dataset$abundance)
```

```{r}
# Get Population values

## Put 0 in monitoring of each species
test <- toy_dataset %>%
  select(siteid, op_id, species, abundance) %>%
  pivot_wider(names_from = "species", values_from = "abundance") %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))

## Imput missing values

## Compute changes by population

## Compute LPI

# GAM Method
#1. fitted a GAM on observed values with log
#10
#(N
#t
#)asthe
#dependent variable and year (t) as the independent,
#2. set the smoothing parameter to the length of the pop-
#ulation time series divided by 2 (Wood 2006),
#3. selected the smoothing-parameter value by comparing
#the estimated degrees of freedom when the smoothing
#parameter was successively incremented by 1,
#4. used fitted GAM values to calculate predicted values
#for all years (including those with no real count data),
#and
#5. averaged and aggregated d values from the imputed
#counts as described earlier
```




