---
title: "Community structure"
author: "Alain Danet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
source(here::here("start_rmd.R"))
```
```{r}
library(vegan)
```


```{r load-targets, include=FALSE}
```

# Diversity

## Hill numbers

### Test on the BCI dataset

```{r}
data(BCI)
H <- diversity(BCI)
simp <- diversity(BCI, "simpson")
invsimp <- diversity(BCI, "inv")
bci_div <- tibble(
  species_nb = specnumber(BCI),
  shannon = H,
  simp = simp,
  inv_simp = invsimp,
  evenness = H / log(species_nb)
)
```

```{r}
kable(bci_div)
```


### Test on my dataset

```{r}
tar_load(c(hillnb, chao_hillnb))

colnames(chao_hillnb) <- str_replace(colnames(chao_hillnb), "chao_", "")
chao_hillnb <- chao_hillnb %>%
  mutate(type = "chao") %>%
  select(op_id, richness:last_col())

chao_hillnb %>%
  filter(is.na(evenness)) %>%
  mutate(evenness2 = shannon / log(richness))

hillnb <- hillnb %>%
  rename(richness = species_nb) %>%
  mutate(type = "classic") %>%
  select(op_id, type, richness, shannon, simpson, evenness)

div <- rbind(chao_hillnb, hillnb)
```

```{r}
div_summary <- div %>%
  group_by(type) %>%
  summarise(
    across(
      where(is.double),
      ~list(
        enframe(
          summary_distribution(x = .x)
        )
      )
    )
    ) %>%
  pivot_longer(cols = c(richness:evenness), names_to = "facet", values_to =
    "biodiversity") %>%
  unnest(cols = "biodiversity") %>%
  pivot_wider(names_from = "name", values_from = "value")
```

```{r}
div_summary %>%
  kable(.,
    caption = "Summary of diversity indices distribution",
    label = "tab:div-comp"
  )
```

- Chao indices leads to higher values of evenness, shannon and simpson, in
  average and also to more spread values 

```{r}
div %>%
  pivot_longer(cols = c(richness:evenness), names_to = "facet", values_to =
    "diversity") %>%
ggplot(aes(y = diversity, x = type)) +
geom_violin() +
facet_grid(rows = vars(facet), scales = "free_y")
```

- Look at infinit values (now fixed, it was for richness == 1)

- Computed on chao estimations of shannon and richness, most values pielou evenness are
  above one:

```{r}
chao_hillnb %>%
  filter(evenness > 1) %>%
  kable
```

### Relationships between Chao and classic number

- Species richness: in Chao we see interpolation and extrapolation to put everyone on
  the level (exemple: community with 70 species had 55 species equivalent
  according to Chao).

```{r}
div <- div %>%
  arrange(op_id, type)
stopifnot(all(div[div$type == "chao", ]$op_id == div[div$type == "classic", ]$op_id))
```

```{r}
old_par <- par()
par(mfrow = c(2, 2))
for (i in c("richness", "shannon", "simpson", "evenness")) {
  plot(
    div[div$type == "chao", ][[i]] ~ div[div$type == "classic", ][[i]],
    ylab = "Chao",
    xlab = "Classic"
  )
  abline(1, 1)
  title(i)
}
par(old_par)
```

## Chao

### Test 

```{r, eval=FALSE}
test <- filtered_dataset$measurement %>%
  filter(op_id %in% unique(op_id)[1:2])
ti <- get_chao_hillnb(
  x = test,
  coverage = .985,
  confidence_int = NULL)
```

# Vegdist turnover

```{r}
tar_load(c(com_mat_site, rigal_trends, var_temporal_trends))
get_vegdist_temporal_turnover(com_mat_site$mat[[1]], method = "jaccard")
names(rigal_trends) <- var_temporal_trends
trend_vegdist <- rigal_trends[names(rigal_trends) %in% c("jaccard", "horn", "chao")]
trend_vegdist_slope <- map_dfr(trend_vegdist,
  ~.x[, c("siteid", "linear_slope")], .id = "response")
```

```{r}
trend_vegdist_slope %>%
  ggplot(aes(x = linear_slope)) +
  geom_histogram() +
  facet_grid(cols = vars(response)) +
  labs(y = "Count", x = "Slope of similarity vs year")
```


## Analysis

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>
