---
title: "RivFishTIME coverage"
author: "Alain Danet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    df_print: paged
    toc: true
    toc_float: true
    theme: "readable"
    code_folding: hide
bibliography: "bibliography.bib"
---

```{r setup, include=FALSE}
source(here::here("start_rmd.R"))
```

# Outline 

- Biodiversity trends are assessed with available timeseries
  - Biaised toward western countries
  - How well do they cover global change?
- What is the environmental coverage of usual timeseries dataset?
  - Crucial when assessing trends without taking in account drivers because in
    that case the value of your conclusion can only be extrapolated on the
    critical assumption that the analyzed dataset is representative of the
    considered statistical population
  - Here, we use a dataset dominated by western countries 
    - Trends in streamflow, water temperature, changes in anthropogenic
      pressures
    - Coverage in land-use


# Land-use and stream characteristics coverage

```{r, eval = FALSE}
tar_load(c(riveratlas_total, snapped_site_river, riveratlas_site))
evt_riv <- c("tmp_dc_cyr", "hft_ix_c93", "dist_up_km", "ord_stra", "dis_m3_pyr", "ele_mt_cav", "pac_pc_cse", "urb_pc_cse", "hft9309_perc")
match_site_riveratlas <- do.call(rbind, snapped_site_river[!map_lgl(snapped_site_river, ~all(is.na(.x)))]) %>%
  filter(!is.na(riverid))

riveratlas_total <- riveratlas_total %>%
  mutate(
    rivfishtime = ifelse(hyriv_id %in% match_site_riveratlas$riverid, TRUE, FALSE),
    hft9309_perc = ((hft_ix_c09 - hft_ix_c93) / hft_ix_c93) * 100
    ) %>%
  select(-hft_ix_c09) %>%
  mutate(across(c(slp_dg_cav, tmp_dc_cyr, hft_ix_c93), ~.x / 10))
```
```{r, eval = FALSE}
riveratlas_total_bp_stat <- riveratlas_total %>%
  group_by(rivfishtime) %>%
  summarise(across(!c(hyriv_id), function(x) {
      stats <- boxplot.stats(x)$stats
      list(data.frame(
          min = stats[1], q25 = stats[2], median = stats[3],
          q75 = stats[4], max = stats[5]
          ))
  })) %>%
  pivot_longer(!c(rivfishtime), names_to = "variable", values_to = "values") %>%
  unnest(values)
save(riveratlas_total_bp_stat, file = here("data", "riveratlas_total_bp_stat.rda"))
rm(riveratlas_total, riveratlas_site)
gc()
```

- Anthropogenic pressures:
  - RivFishTIME sites are located in areas much more degraded than globally
    (median human footprint index 1993 three times higher)
  - Paradoxally, RivFishTIME sites are also located where anthropogenic
    pressures decreased in average across the last decades (more than 50% of the sites),
    while at global scale we observe recent increase in anthropogenic pressures
    (75% of the sites had a increase from +0% up to +75%).
  - Linked with anthropogenic pressures: higher percentage of protected, urban, cropland and pasture
    areas, and higher population density than globally 


```{r, fig.width = 10, fig.asp = 2.5}
load(file = here("data", "riveratlas_total_bp_stat.rda"))
riveratlas_var_name <- c(
  get_rev_vec_name_val(c(
      get_river_atlas_significant_var(),
      get_land_class_var())),
  "hft9309_perc" =  "Human footprint change 93-09 (%)"
)
riveratlas_total_bp_stat %>%
  mutate(variable = riveratlas_var_name[variable]) %>%
  ggplot(
    aes(x = rivfishtime,
      lower = q25, upper = q75, middle = median, ymin = min, ymax = max)
    ) +
  geom_boxplot(stat = "identity") +
  scale_x_discrete(breaks = c(FALSE, TRUE), labels = c("RiverAtlas", "RivFishTIME")) +
  facet_wrap(vars(variable), scale = "free", ncol = 3) +
  theme(axis.title.x = element_blank())
```

# Temporal changes in temperature and streamflow 

- TODO:
  - Futurestreams:
    - [ ] Build monthly average SpatRast (1969-2005)
    - [ ] Build annual max monthly SpatRast [@comte_climate_2021]
    - [ ] Get crap lm trends over SpatRast
    - [ ] redo with one GCM model?

- Methods:
  - Take the temp max of each avg month value 
  - Linear model

```{r, eval=TRUE, fig.cap="Use of previous water temperature dataset"}
tar_load(c(wt_mv_avg_roll, wt, site_desc_loc))

tu <- wt_mv_avg_roll %>%
  mutate(year = year(date)) %>%
  group_by(siteid, year) %>%
  summarise(
    max_m_tmp = max(mv_avg_12m),
    med_m_tmp = median(mv_avg_12m),
    .groups = "drop"
  )

ml <- glmmTMB(max_m_tmp ~ year + (1 + year | siteid), tu)
tibble(y = coef(ml)$cond$siteid[, "year"] * 10) %>%
  ggplot(aes(x = y)) +
  geom_histogram(bins = 100) +
  labs(
    x = "Annual maximum water temperature trends (°C by decade) RivFishTIME",
    y = "Number of sites"
  )
```

```{r, eval = FALSE}
summary_distribution(coef(ml)$cond$siteid[, "year"] * 10)
str(coef(ml))

inla(
  tmp ~ date +
  f(siteid, model = 'iid') +
  f(siteid_year, date, model = 'iid'),
data = wt_mv_avg_roll %>%
  filter(siteid %in% sample(siteid, 100)) %>%
  mutate(siteid_year = siteid)
)
```



# Misc {-}

## Analysis

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
if(requireNamespace('git2r', quietly = TRUE)) {
  git2r::repository()
} else {
  c(
    system2("git", args = c("log", "--name-status", "-1"), stdout = TRUE),
    system2("git", args = c("remote", "-v"), stdout = TRUE)
  )
}

## session info
sessionInfo()
```

</details>

# References {-}
