---
title: "Main text"
author: "Alain Danet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2
bibliography: "bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 1.618, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "t",  # pdf mode
                      fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300,
                      out.width = "100%",
                      dev = "svg",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")

library(here)
source(here::here("R", "misc.R"))
source(here::here("R", "variable_shortcut.R"))
source(here::here("R", "basic_stat.R"))
library(tidyverse)
library(magrittr)
library(cowplot)
library(targets)

if (Sys.info()["login"] == "alain")
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib")
  )
```

```{r load-targets, include=FALSE}
tar_load(c(site_desc_loc, measurement))
tar_load(c(modelling_data, filtered_dataset_modelling, analysis_dataset))
tar_load(filtered_op_protocol_modelling)
tar_load(c(measurement_exo, modelling_data_exo))

measurement_exo <- measurement_exo %>%
  filter(op_id %in% unique(modelling_data_exo$op_id))

tar_load(c(
    gaussian_inla_no_drivers_effects,
    gaussian_inla_exo_no_drivers_effects)
)
inla_no_drivers_effects <- rbind(
  gaussian_inla_no_drivers_effects,
  gaussian_inla_exo_no_drivers_effects
)

tar_load(c(pred_number, pred_inla))
```
```{r}
table_to_vec <- function(x) {
  setNames(as.numeric(x), names(x))
}
get_comp <- function(
  pred = pred_number,
  type = "pred_hft93_t0",
  resp = NULL,
  lvl_comp = NULL
  ) {
  pred[[type]][[resp]][["mean"]][lvl_comp]
}
```

- Article guidelines by Nature in Ecology and Evolution: 
  - Main text – up to 3,500 words, excluding abstract, Methods, references and figure legends.
  - Abstract – up to 200 words, unreferenced. 
  - Display items – up to 6 items (figures and/or tables). 
  - Extended Data – up to 10 items (figures and/or tables, linked from the main text in the html version of the paper).
  - Article should be divided as follows: 
    - Introduction (without heading, up to 500 words of referenced text on the background of the work) 
    - Results (concise, focused account of the findings)
    - Discussion
    - Online Methods.
- Results and Methods should be divided by topical subheadings; the Discussion does not contain subheadings.
- References – as a guideline, we typically recommend up to 50.


# Introduction (up to 500 words)

### 1.1 Community changes

- Biodiversity crisis?
  - IPBES
  - Report on population declines (LPI) 
    - LPI for freshwater population declines by 83% between 1970 and 2014
- Community changes crisis?
  - Contrasted results in temporal trends:
    - High change in community change but no change in average in species
      richness [@vellend_global_2013; @dornelas_assemblage_2014;
      @blowes_geography_2019], or even positive in aquatic insects
      [@van_klink_meta-analysis_2020]

### 1.2 Antropogenic drivers

- Land-use & Over-exploitation are the main drivers of biodiversity losses (IPBES,
  2019; @reid_emerging_2019)
  - Spatial comparison: @newbold_global_2015 (terrestrial vertebrates),
    @outhwaite_agriculture_2022 (terrestrial insects)
  - Land-use in freshwater system is a major and continuous threats [@reid_emerging_2019], 
    including pollution by cropland and urban zones.
  - examples of land-use on freshwater biodiversity
    - pollution by coal mines: @giam_impact_2018
    - Ref in @reid_emerging_2019
    - Flow alteration by dams decreased discharge, impedes fish migration, and
      decrease the connectivity between terrestrial and aquatic compartment
      [@juracek_aging_2015].
    - Non-native fish species introduction can alter native fish abundance but
      also invertebrates and macrophytes [@gallardo_global_2016]

### 1.3 Biodiversity facets

- Trends in species richness, abundance, in dissimilarity
  [@dornelas_assemblage_2014; @blowes_geography_2019; @lyashevska_how_2012]
- Contrasted results and patterns
- Do community trends of different biodiversity facets lie in different
  dimensions? Or are they linear combination of each other? [@lyashevska_how_2012]

### 1.4 Freshwater Fishes

- Research gap on aquatic freshwater fishes
- Critical elements of freshwater ecosystems
- Big anthropogenic pressures: receptacle of terrestrial land alteration + flow
  alteration + pollution + introduction of non-native species + exploitation
  [@reid_emerging_2019]

### 1.5 Study objectives

- Assess the effect of anthropogenic pressures on the spatio-temporal trends of
  fish freshwater communities at global scale
- Assess the dimensionality of community changes?

# Results

```{r}
get_global_effect <- function (
  effect = inla_no_drivers_effects,
  resp = NULL,
  ci_lvl = "level:0.95"
  ) {

  out <- effect %>% 
    filter(
      ci_level == ci_lvl,
      term == "log1_year_nb",
      response == resp 
    )

  out[, c("low", "high", "mean")] %>%
    pivot_longer(everything()) %>%
    deframe()
}

p_ci <- function(x, r = 2, p = TRUE) {
  out <- round(x, r)

  if (p) {
    paste0(out["mean"], "%", " [", out["low"],"%,", out["high"],"%]")
  } else {
    paste0(out["mean"], " [", out["low"],",", out["high"],"]")
  }

}

get_effect_ci <- function(
  effect = tu,
  resp = NULL,
  term = "log1_year_nb",
  ci_lvl = "level:0.95",
  r = 2
  ) {
  term1 <- term

  out <- effect %>%
    filter(
      ci_level == ci_lvl,
      term == term1,
      response == resp
    )

  out <- out[, c("low", "high", "mean")] %>%
    pivot_longer(everything()) %>%
    deframe()

  p_ci(x = out, r = r, p = FALSE)

}
```

```{r}
trends <- map(unique(inla_no_drivers_effects$response),
  ~get_global_effect(
  effect = inla_no_drivers_effects,
  resp = .x,
  ci_lvl = "level:0.95"
  )
)
names(trends) <- unique(inla_no_drivers_effects$response)

trends80 <- map(unique(inla_no_drivers_effects$response),
  ~get_global_effect(
  effect = inla_no_drivers_effects,
  resp = .x,
  ci_lvl = "level:0.80"
  )
)
names(trends80) <- unique(inla_no_drivers_effects$response)

trends90 <- map(unique(inla_no_drivers_effects$response),
  ~get_global_effect(
  effect = inla_no_drivers_effects,
  resp = .x,
  ci_lvl = "level:0.90"
  )
)
names(trends90) <- unique(inla_no_drivers_effects$response)

```

```{r}
abun_trends <- p_ci(
  log_beta_to_perc_rate(trends$log_total_abundance)
  * log(10 + 1),
  r = 1)
rich_trends <- p_ci(
  log_beta_to_perc_rate(trends$log_chao_richness)
  * log(10 + 1),
  r = 1
)
hill_trends <- p_ci(
  trends$hillebrand_dis_scaled
  * log(10 + 1),
  r = 2, p = FALSE
)
exo_trends <- map_chr(c("perc_exo_sp", "perc_exo_abun"),
  ~p_ci(trends[[.x]] * log(10 + 1), r = 3, p = FALSE)
)
names(exo_trends) <- c("perc_exo_sp", "perc_exo_abun")

exo_trends90 <- map_chr(c("perc_exo_sp", "perc_exo_abun"),
  ~p_ci(trends90[[.x]] * log(10 + 1), r = 4, p = FALSE)
)
names(exo_trends90) <- c("perc_exo_sp", "perc_exo_abun")

exo_trends80 <- map_chr(c("perc_exo_sp", "perc_exo_abun"),
  ~p_ci(trends80[[.x]] * log(10 + 1), r = 3, p = FALSE)
)
names(exo_trends80) <- c("perc_exo_sp", "perc_exo_abun")
```

```{r}
extract_pred <- function (x = NULL, resp = NULL, val_to_extract = NULL) {

  ci_low <- x[[resp]]$quant0.025[val_to_extract]
  ci_high <- x[[resp]]$quant0.975[val_to_extract]

  # Does the ci at 95 overlap? 
  ci_overlap <- ci_low[val_to_extract[1]] > ci_high[val_to_extract[2]] |
    ci_high[val_to_extract[1]] < ci_low[val_to_extract[2]]

  list(
    val = x[[resp]]$mean[val_to_extract],
    ci_overlap = !ci_overlap
  )
}
#' Get percentage of difference, the first value being the reference
get_ratio <- function(x, r = 1) {
  #reduce(x, `/`)
  out <- round((x[2] - x[1]) / x[1] * 100, r)
  paste0(out, "%")
}
```

```{r}
tar_load(pred_number)
```

```{r}
pred_hft93_t0 <- map(
  setNames(names(pred_number$pred_hft93_t0), names(pred_number$pred_hft93_t0)),
  ~extract_pred(
    x = pred_number$pred_hft93_t0,
    resp = .x,
    val_to_extract = c("intact", "max")
  )
)

abun_rich_hft93_int_max <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft93_t0[[.x]]$val), r = 0)
)
other_hft93_int_max <- map_chr(
  setNames(names(pred_number$pred_hft93_t0), names(pred_number$pred_hft93_t0))[
    !names(pred_number$pred_hft93_t0) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft93_t0[[.x]]$val, r = 0)
  )
```


```{r}
pred_hft9309_t0 <- map(
  setNames(names(pred_number$pred_hft9309_t0), names(pred_number$pred_hft9309_t0)),
  ~extract_pred(
    x = pred_number$pred_hft9309_t0,
    resp = .x,
    val_to_extract = c("div/2", "inc/2")
  )
)

abun_rich_hft93909_dec_inc_by2 <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft9309_t0[[.x]]$val), r = 0)
)
other_hft93909_dec_inc_by2 <- map_chr(
  setNames(names(pred_number$pred_hft9309_t0), names(pred_number$pred_hft9309_t0))[
    !names(pred_number$pred_hft9309_t0) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft9309_t0[[.x]]$val, r = 0)
  )
```

```{r}
pred_hft93_t10 <- map(
  setNames(names(pred_number$pred_hft93_t10), names(pred_number$pred_hft93_t10)),
  ~extract_pred(
    x = pred_number$pred_hft93_t10,
    resp = .x,
    val_to_extract = c("intact", "max")
  )
)

abun_rich_hft93_int_max_t10 <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft93_t10[[.x]]$val), r = 0)
)
other_hft93_int_max_t10 <- map_chr(
  setNames(names(pred_number$pred_hft93_t10), names(pred_number$pred_hft93_t10))[
    !names(pred_number$pred_hft93_t10) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft93_t10[[.x]]$val, r = 0)
  )
```

```{r}
pred_hft9309_t10 <- map(
  setNames(names(pred_number$pred_hft9309_t10), names(pred_number$pred_hft9309_t10)),
  ~extract_pred(
    x = pred_number$pred_hft9309_t10,
    resp = .x,
    val_to_extract = c("div/2", "inc/2")
  )
)

abun_rich_hft93909_dec_inc_by2_t10 <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft9309_t10[[.x]]$val), r = 0)
)
other_hft93909_dec_inc_by2_t10 <- map_chr(
  setNames(names(pred_number$pred_hft9309_t10), names(pred_number$pred_hft9309_t10))[
    !names(pred_number$pred_hft9309_t10) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft9309_t10[[.x]]$val, r = 0)
  )
```

```{r}
tar_load(c(gaussian_inla_std_effects, gaussian_inla_exo_std_effects))
tu <- rbind(gaussian_inla_std_effects, gaussian_inla_exo_std_effects)
ci80 <- tu %>%
  filter(ci_level == "level:0.80")
ci90 <- tu %>%
  filter(ci_level == "level:0.90")
ci95 <- tu %>%
  filter(ci_level == "level:0.95")
```

## Global temporal trends

We found strong evidence for a global increase in both total abundance and
species richness of freshwater fish communities, respectively `r abun_trends`
and `r rich_trends` by decade (CI 95%). We further found moderate evidence for a
decline in the proportion of non-native abundance,
`r exo_trends[["perc_exo_abun"]]` by decade (CI 90%) but no evidence for a net
temporal trend in the proportion of non-native richness 
(`r exo_trends[["perc_exo_sp"]]`, CI 80%). We also found that temporal
dissimilarity metrics of Jaccard, Turnover and Simpson-based have overall effect
comparable in magnitude (Fig. S10), indicating the temporal changes of fish communities
are driven by both species replacement and changes in species dominance.

## Effect of anthropogenic pressures 

We found strong evidence that past anthopogenic pressures, quantified by the
human footprint index at the year 1993, result in higher species richness (Fig.
1), as most degraded sites have `r abun_rich_hft93_int_max["log_chao_richness"]`
more richness than intact sites. Sites with higher past anthropogenic pressures
also had an higher proportion of non-native richness and non-native abundance
(strong evidence, Fig. 1), such as most degraded sites had a proportion of
non-native species twice higher than intact sites (resp.  
`r scales::percent(pred_hft93_t0$perc_exo_sp$val["max"])` and 
`r scales::percent(pred_hft93_t0$perc_exo_sp$val["intact"])`) and a proportion
of non-native abundance three times higher 
(`r scales::percent(pred_hft93_t0$perc_exo_abun$val["max"])` and
`r scales::percent(pred_hft93_t0$perc_exo_abun$val["intact"])` resp.). 
We found also strong evidence that past anthropogenic pressures resulted in
lower total abundance (Fig. 1), such as most degraded sites had 
`r abun_rich_hft93_int_max["log_total_abundance"]` lower total abundance than
intact sites.

```{r}
hft93abun <- get_effect_ci(resp = "log_total_abundance", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
hft93rich <- get_effect_ci(resp = "log_chao_richness", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
hft93exor <- get_effect_ci(resp = "perc_exo_sp", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c93")
hft93exoa <- get_effect_ci(resp = "perc_exo_abun", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c93")
hft93hill <- get_effect_ci(resp = "hillebrand_dis_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
hft93turn <- get_effect_ci(resp = "turnover_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
```
We found strong evidence that past anthropogenic pressures increased the
temporal trends of total abundancen, species richness (resp. Credible Interval 95%:$\beta\prime =$ 
`r hft93abun` and `r hft93rich`, Fig. 1), and we found weak evidence that past
anthropogenic pressures increased the temporal trends of temporal trends of
the proportion of non-native richness (CI 80%:$\beta\prime =$ `r hft93exor`,
Fig. 1). However we found no evidence that past anthropogenic pressures affect
the proportion of non-native abundance (CI 80%:$\beta\prime =$ `r hft93exoa`,
Fig. 1). We also found strong evidence that past anthropogenic pressures
increased the temporal trends of temporal dissimilarity and turnover, i.e.
respectively indicating changes in species dominance and species replacement
(resp. CI 95%:$\beta\prime =$ `r hft93hill` and `r hft93turn`).

In contrast, we found strong evidence that the recent increase in anthropogenic pressures,
quantified by the ratio of the human footprint index between 2009 and 1993,
decreased both species richness and total abundance (Fig. 1). 
A multiplication by two of human footprint between 1993 and 2009 decreased
species richness by `r abun_rich_hft93909_dec_inc_by2["log_chao_richness"]` 
and total abundance by `r abun_rich_hft93909_dec_inc_by2["log_total_abundance"]`
compared to a division by two of human footprint in the same time range. We also
found strong evidence that recent increase in anthropogenic pressures increased
the proportion of non-native species and abundances (Fig. 1), respectively by five times
(`r scales::percent(pred_hft9309_t0$perc_exo_sp$val["div/2"])` vs
`r scales::percent(pred_hft9309_t0$perc_exo_sp$val["inc/2"])` for sites whose
anthropogenic pressures recently decrease and increase respectively) and by
eight times (`r scales::percent(pred_hft9309_t0$perc_exo_abun$val["div/2"])` vs
`r scales::percent(pred_hft9309_t0$perc_exo_abun$val["inc/2"])`).

```{r}
hft09abun <- get_effect_ci(resp = "log_total_abundance", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c9309_log2_ratio", r = 3)
hft09rich <- get_effect_ci(resp = "log_chao_richness", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c9309_log2_ratio", r = 3)
hft09exor <- get_effect_ci(resp = "perc_exo_sp", ci_lvl = "level:0.90", term = "log1_year_nb:hft_ix_c9309_log2_ratio", r = 2)
hft09exoa <- get_effect_ci(resp = "perc_exo_abun", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c9309_log2_ratio", )
hft09hill <- get_effect_ci(resp = "hillebrand_dis_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c9309_log2_ratio")
hft09turn <- get_effect_ci(resp = "turnover_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c9309_log2_ratio")
```

We did not found evidence for an effect of recent anthropogenic
pressures on the temporal trends of both total abundance and species richness
(resp. CI 80%: $\beta\prime =$ `r hft09abun` and `r hft09rich`, Fig. 1). However, we found
moderate evidence that recent increase in anthropogenic pressures increase the
temporal trends of the proportion in non-native richness (CI 90%: $\beta\prime =$ `r hft09exor`,
Fig. 1) but we found no evidence for an effect on the temporal trends of the
proportion of non-native abundance (CI 80%: $\beta\prime =$ `r hft09exoa`, Fig. 1). We found
also strong evidence that recent increase in anthropogenic pressures resulted in
stronger species turnover and dissimilarity (resp. CI 95%: $\beta\prime =$ `r hft09turn` and
`r hft09hill`, Fig. 1).

```{r}
hft2abun <- get_effect_ci(resp = "log_total_abundance", ci_lvl = "level:0.95",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio", r = 3)
hft2rich <- get_effect_ci(resp = "log_chao_richness", ci_lvl = "level:0.80",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio", r = 3)
hft2exor <- get_effect_ci(resp = "perc_exo_sp", ci_lvl = "level:0.90",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio", r = 3)
hft2exoa <- get_effect_ci(resp = "perc_exo_abun", ci_lvl = "level:0.80",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio")
hft2hill <- get_effect_ci(resp = "hillebrand_dis_scaled", ci_lvl = "level:0.95",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio")
hft2turn <- get_effect_ci(resp = "turnover_scaled", ci_lvl = "level:0.95",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio")
```

We also found strong evidence for antagonistic effect between past and recent
anthropogenic pressures on the temporal trends of total abundance
(CI 95%: $\beta\prime =$ `r hft2abun`, Fig.1), meaning that recent anthropogenic decreased more
temporal trends of total abundance in sites that were previously more degraded.
We did not find evidence for such interaction for species richness
(CI 80%: $\beta\prime =$ `r hft2rich`, Fig. 1). We also found moderate evidence for a
synergistic effect between past and recent anthropogenic pressures on the
temporal trends of the proportion of non-native species, but not of non-native
abundance (resp. CI 90%: $\beta\prime =$ `r hft2exor` and CI 80% `r hft2exoa`, Fig.1). Finally, we
found strong evidence for a synergistic between past and recent anthropogenic
pressures on temporal dissimilarity and turnover (resp. CI 95%: $\beta\prime =$ `r hft2hill` and
`r hft2turn`, Fig.1).

## Typology of community temporal trends

We found that the Principal Component Analysis necessitated four principal
components over the four community metrics to explain more than 90% of variation
in temporal trends (69% and 87% with respectively two and three principal
components), one axis for each community metrics, meaning that there was a low
collinearity among the temporal trends of community metrics (Fig. 2). Clustering
of commmunity temporal trends identified six clusters related to PCA axis.
Around 40% of the sites were attributed to a cluster related to a medium temporal increase
in species richness and total abundance, as well a medium temporal dissimilarity
and turnover (Fig. 2). The second cluster (16% of the sites) was related to
communities displaying an high turnover, and a relatively moderate temporal
increase in species richness, total abundance and dissimilarity. The third, fourth
and fifth clusters (resp. 13%, 12% and 9% of the sites, Fig. 2) were
respectively related to a temporal increase in species richness, temporal
decrease in total abundance and to a temporal decrease in species richness. The
last cluster (7% of the sites) was related to communities displaying relatively
low temporal changes.

## Spatial distribution of community temporal trends 

We found that the spatial distribution of clusters of community temporal trends
were relatively homogeneous across realms and more heterogeneous across
hydrographic basin (Fig. 3). We further found that the random variance in
temporal trends, i.e. in the slope coefficient of time predictor, attributed to
site identity was higher than the one attributed to hydrographic basin for all
variables (Table S14). Expect for total abundance and species richness, the
variance attributed to site was at least twice higher than the one attributed to
basin (Table S14).


# Discussion

### Global trends

- We found the same temporal trends in total abundance than
  @van_klink_meta-analysis_2020 (+11%) in aquatic insects.
- Turnover is the main components of dissimilarity as found in previous studies @blowes_geography_2019
- We further found that dissimilarity is both related to change in dominance and
  in species occurrence, i.e. changes in changes in species occurrence concerns
  abundant species [@hillebrand_biodiversity_2018]
- Our results on species richness contrast with @dornelas_assemblage_2014

### Effect of anthropogenic pressures

- Both past, recent pressures and their interactions increase temporal dissimilarity
- We observe stronger recovery in species richness and abundance in sites more degraded 30 years ago
  - Might be due to DCE in Europe and Clean Water Act (US), (are there same
    thing in Canada, Australia and JPN?)
  - However, 92% of our sites can be categorized as degraded
  - Might explain @van_klink_meta-analysis_2020
- Recent increase in human pressures left a strong spatial signature on richness
  and abundance, confirming previous results on terrestrial vertebrates and
  insects [@newbold_global_2015; @outhwaite_agriculture_2022]
  - but was found in the temporal trends only on abundance in
    interaction with past pressures, i.e. recent increases in anthropogenic
    pressures seems to have an effect in most degraded sites in the past.
- Anthropogenic pressures have strong spatial and temporal effects on community
  changes

### Constrasted community changes

- Our results shows that community changes in different dimensions: richness,
  abundance and dissimilarity, in accordance with previous results
  [@dornelas_assemblage_2014; @blowes_geography_2019]
- Community can change in different ways: richness, abundance, dissimilarity.
  Related to the multidimensional nature of biodiversity [@lyashevska_how_2012;
  @chase_embracing_2018]
- Clustering reveals that most of our sites displays medium changes in
  dissimilarity, low increase in richness and in abundance, to relate with the fact that
  92% of our sites can be considered as degraded and that the model suggests
  that past degradation increases the temporal trends of richness and abundance

### Spatial scale of community changes (minor point)

- Variance of temporal trends attributed to site twice higher than basin
- Cluster distribution of community changes are homogeneous across realms,
  relatively homogeneous across basin
- Spatial heterogeneity of community changes at small spatial scale, echoes with 
  abundant research on richness changes across spatial scale [@chase_embracing_2018]
  Our results suggest that the way that community changes is very heterogeneous
  at small spatial scale and might reflect the spatial heterogeneity in
  anthropogenic pressures and stream characteristics.

## 5 Conclusion

- Temporal trends confirm previous results on dissimilarity and species richness
  [@van_klink_meta-analysis_2020; @dornelas_assemblage_2014;
  @blowes_geography_2019]
- Past and recent changes in anthropogenic pressures have strong effects on
  temporal trends, in addition to spatial signature [@newbold_global_2015;
  @outhwaite_agriculture_2022]
- Contrasted community changes: either species richness, abundance,
  dissimilarity

# Wordstack

# References {-}

<div id="refs"></div>
