---
title: "Past and recent anthropogenic pressures drive community changes over space and time in riverine fish assemblages"
author: "Alain Danet, Xingli Giam, Julian Olden, Lise Comte"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \hypersetup{backref=true}
output:
  bookdown::pdf_document2:
bibliography: "bibliography.bib"
csl: nature.csl
link-citations: true
indent: true
spacing: double
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 1.618, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "t",  # pdf mode
                      fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300,
                      out.width = "100%",
                      dev = "svg",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")

library(here)
source(here::here("R", "misc.R"))
source(here::here("R", "variable_shortcut.R"))
source(here::here("R", "basic_stat.R"))
library(tidyverse)
library(magrittr)
library(cowplot)
library(targets)
library(kableExtra)


if (Sys.info()["login"] == "alain")
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib")
  )
```

```{r load-targets, include=FALSE}
tar_load(c(site_desc_loc, measurement))
tar_load(c(modelling_data, filtered_dataset_modelling, analysis_dataset))
tar_load(filtered_op_protocol_modelling)
tar_load(c(measurement_exo, modelling_data_exo))


fnsite <- nrow(filtered_dataset_modelling$location)

measurement_exo <- measurement_exo %>%
  filter(op_id %in% unique(modelling_data_exo$op_id))

tar_load(c(
    gaussian_inla_no_drivers_effects,
    gaussian_inla_exo_no_drivers_effects)
)
inla_no_drivers_effects <- rbind(
  gaussian_inla_no_drivers_effects,
  gaussian_inla_exo_no_drivers_effects
)

tar_load(c(pred_number, pred_inla))
```
```{r}
table_to_vec <- function(x) {
  setNames(as.numeric(x), names(x))
}
get_comp <- function(
  pred = pred_number,
  type = "pred_hft93_t0",
  resp = NULL,
  lvl_comp = NULL
  ) {
  pred[[type]][[resp]][["mean"]][lvl_comp]
}
```

- TODO: 
  - [x] GLMMTMB verifier R2: function does not work on glmmTMB object but I checked
    the function and the formula is the same that I used
  - [x] human footprint index / non-native richness
  - [x] julian role of native in degraded ecosytems: tolerant non-native 
  - [x] limitation of the study: climate change
  - [ ] Is it possible to capture the number of spliting of fish species in the
    database? (if not available, may be prepare a mail for fishbase)
  - [ ] Supplementary
    - [ ] Grosse table avec toutes les variables, main text names, les distributions
    - [ ] refaire graph correspondance données / dates des human footprint index
    - [ ] Mapper les tendances de biodiversité plus histogrammes
    - [ ] Review all comments
  - [ ] Review figures:
    - [x] Check PCA axis contribution: it is the varimax rotation of the PC that
      equalized the contribution of each axis. 
      - [ ] Replace by non rotated PCA? 
    - [ ] Figure 
      - [ ] fig 3: enlever l'antartique, agrandir la carte du monde, enlever les pays
      dans les labels des histrogrammes des clusters, forme des insets doit être
      égal à la forme des bassins, ordre des bassins dans histogrammes doit
      correspondre à celui des insets
      - [ ] Review all figure comments 
  - [ ] Table random variance + both R-squared in main text

```{r, results="hide"}
word_count <- wordcountaddin::text_stats(here("paper", "main_text.Rmd"))
readable <- wordcountaddin::readability(filename = here("paper", "main_text.Rmd"), quiet = TRUE)
```
```{r}
word_count
```
- Article guidelines by Nature in Ecology and Evolution: 
  - Main text – up to 3,500 words, excluding abstract, Methods, references and figure legends.
  - Abstract – up to 200 words, unreferenced. 
  - Display items – up to 6 items (figures and/or tables). 
  - Extended Data – up to 10 items (figures and/or tables, linked from the main text in the html version of the paper).
  - Article should be divided as follows: 
    - Introduction (without heading, up to 500 words of referenced text on the background of the work) 
    - Results (concise, focused account of the findings)
    - Discussion
    - Online Methods.
- Results and Methods should be divided by topical subheadings; the Discussion does not contain subheadings.
- References – as a guideline, we typically recommend up to 50.

\pagebreak

# Abstract (up to 200 words; now: 284)

Recent studies found increase of total abundance and occupancy over time in
freshwater insect assemblages, suggesting that these could be the consequence of
recent water quality improvement in industrial countries, while other studies
found no net trends in local species richness but high species turnover in
marine, freshwater and terrestrial assemblages. Antropogenic pressures is one of
the major drivers of biodiversity loss and community composition changes. Here,
we test how past and recent changes in anthropogenic pressures drive community
changes over space and time in fish riverine communities. Here we show that both
past and recent antropogenic pressures result in lower total abundance (-29%
each) while having a proportion of non-native richness and abundance three to
five times higher than most intact sites. Similary to other studies, we found a
global increase in species richness (+xx) and in total abundance (+xx) over time. 
We found that past anthropogenic pressures increase the species richness and
total abundance over time but that recent increases pressures had an opposite
effect, stronger in sites the most previously degraded. We further found that
both past and recent anthropogenic pressures increase, synergistically, the
proportion of non-native richness, species replacement, and changes in dominance
over time. We suggest that increase of species richness and total abundance over
time might be due to the selection of most tolerant species. Our results further
confirmed that the temporal trends of different community metrics are poorly
related with each other and that community changes over time are highly
heterogenenous at small spatial scale. By investigating community changes over
time with multiple metrics, our study provides new line of evidence to interpret
recent results in the light of both past and recent changes in antropogenic
pressures.


# Introduction (up to 500 words)

The response of ecological communities to anthropogenic pressures is a central
topic for biological conservation, and their documentation over space and time
is required to better understand human impacts and adequately respond to them.
Previous studies on community changes showed contrasted results, i.e. no net
change in species richness over time in marine and terrestrial assemblages but
important changes in species composition [@vellend_global_2013;
@dornelas_assemblage_2014; @blowes_geography_2019]. In freshwater systems,
recent studies even showed an increase in insect abundance
[@van_klink_meta-analysis_2020] and occupancy [@outhwaite_complex_2020] since
1990s, which raised concerns about the data quality and representativity
[@cardinale_is_2018; @desquilbet_comment_2020], and about the interpretation of
theses increases over time [see comments to @van_klink_meta-analysis_2020]. It is thus
important to pursue the documentation of community changes in freshwater systems
with the best data available and in order to better understand those changes.

Habitat destruction and over-exploitation have been recognised as the
two main drivers of biodiversity losses across ecosystems [@diaz_global_2019;
@caro_inconvenient_2022; @maxwell_biodiversity_2016]. Meta-analysis showed that 
dense urban and agriculture areas are associated with much lower species
richness and total abundance in assemblages of terrestrial vertebrates
[@newbold_global_2015], as well as agricultural areas on terrestrial insect
assemblages [@outhwaite_agriculture_2022]. The effects of anthropogenic
pressures are even considered to be more dramatic on freshwater systems because
they cover a small area of terrestrial land but are the recipients of the
pollution of terrestrial lands, such as urban and industrial wastes and
agricultural pesticides and nutrients [@reid_emerging_2019]. Further
anthropogenic pressures on freshwater systems include flow alteration, that
impedes fish migration and decrease connectivity between terrestrial and aquatic
compartments [@juracek_aging_2015], stoking and harvest for recreational
fishing, and non-native species introduction that can alter native fish
abundance but also invertebrates and macrophytes [@gallardo_global_2016].

Anthropogenic pressures constitute a persistent and continuous threats on
freshwater systems [@reid_emerging_2019], but they might have decreased in time,
as suggested previously to explain the recent increases of insect abundance
and occupancy over time in freshwater systems [@van_klink_meta-analysis_2020;
@outhwaite_complex_2020; @van_strien_modest_2016; @langford_long-term_2009].
However, a study found no clear evidence of such increase in UK water quality
from 1940s [@whelan_is_2022] despite adoptation of several major EU legislations
on water quality. [In the so-called "Western country", the recognition of the
poor water quality in
the 1960s, followed by the implementation of legislations to reverse the
tendency in the 1970s have improved significantly water quality in many
previously degraded places. Examples of such cases is the program of reduction
of sulfur dioxide and nitrogen dioxide emissions that started in 1970 in Sweden,
the creation of the Environmental Protection Agency in 1970 shortly preceded by
adoption of legislations targeting improvement of water quality in the United
States, water agencies creation in France in 1964 followed by a systematic
monitoring of harmful chemicals, etc...] At the same time, anthropogenic
pressures, quantified by the human footprint index, have increased by at least
20 percent between 1993 and 2009 in 75% of the global ecoregions, mostly in
South America, Equatorial Africa, South and South-East Asia, and Oceania
[@venter_sixteen_2016]. While anthropogenic pressures have decreased recently in
historical wealthiest nations, it has increased in the most recent rising 
economies, where are located most of the biodiversity hotspots and the most
intact ecosystems [@venter_sixteen_2016; @williams_change_2020;
@watson_catastrophic_2016]. So, it is important to consider temporal changes in
anthropogenic pressures, to account both for the legacy of the past and its
recent changes.

Here, we explore the effects of anthropogenic pressures, over space and time, on
fish riverine communities. We used the RivFishTime database which gather more
than ten thousands timeseries of fish communities composition and abundances,
monitored mainly in Western and Northern Europe, Northern Eastern America, South
Eastern Australia and Japan. We selected `r fnsite` timeseries (See Methods), median
baseline and completeness of the time series were respectively 1997 and 50%
(Table SX, Fig. SX). We quantified anthropogenic pressures with the human
footprint index, which includes human population density, land-use, and human
built infrastructures such as roads, railways, electric grids, and navigable
pathways [@venter_global_2016]. Previous studies reported that a higher human
footprint index is related to a higher extinction risk in terrestrial mammals, and
that it better predicts mammal distribution range than environmental predictors
[@di_marco_changes_2018; @di_marco_human_2015]. We then quantified past
anthropogenic pressures as the human footprint index of 1993, only 7% of the
riverine fish monitoring in our dataset being realized before this date. Recent changes in
anthropogenic pressures were quantified as the ratio between human footprint
index of 2009 and 1993, enabling us to separate the effects of the legacy of
past anthropogenic pressures and the effects of their recent changes. We
furthermore controlled for the stream gradient, i.e. from upstream to downstream,
which constitutes a major driver of the composition of fish riverine communities
[@reid_emerging_2019, ref] and can be spatially correlated with anthropogenic
pressures [@kuczynski_concomitant_2018; @meybeck_man_1998]. Because biodiversity
is a multidimensional concept whose trends were shown to be not correlated
[@dornelas_assemblage_2014; @blowes_geography_2019; @lyashevska_how_2012;
@hillebrand_biodiversity_2018], we characterised community by multiple metrics,
including species richness and total abundance. We further computed temporal
dissimilarity based on the Simpson diversity index, i.e. based on relative
abundance, which quantify changes in dominance structure in the community
[@hillebrand_biodiversity_2018]. We also computed temporal dissimilarity based
on Jaccard index and its decomposition in turnover, i.e. species replacement,
and nestedness, i.e. species addition or deletion from a common species pool.
The proportion of non-native species in fish communities was also computed in
terms of richness and abundance, based mainly (95% of the species occurrences) on a
published database of species origin status at the basin scale [@tedesco_global_2017].

We hypothesized that high past anthropogenic pressures should result in stronger
increase in species richness and total abundance in fish riverine communities,
i.e. there is a recovery from past anthropogenic pressures, as suggested in
previous studies in aquatic systems [@outhwaite_complex_2020;
@van_klink_meta-analysis_2020; @langford_long-term_2009]. At the same time, we
expected the recovery to be accompanied by an increase in the proportion of
non-native species because more anthroposed areas are related to more propagule
pressure due to human activities [@leprieur_fish_2008; @bernery_freshwater_2022]
and because non-native species are often ubiquitous species [ref]. Recent
increases in anthropogenic pressures were hypothesized to have an opposite
effect of the "recovery" effect from past perturbations on temporal trends of
species richness and total abundance, while it should also increase the
proportion of non-native species. Both anthropogenic pressures are hypothesized
to have a spatial, i.e. a baseline, negative effect on species richness and
total abundance but a positive effect on the proportion of non-native species.
Furthermore, we expected that both anthropogenic pressures interact with each
other, i.e. that the effect of recent increases in anthropogenic pressures is
dependent on the past anthropogenic pressures. In line with previous studies
[@dornelas_assemblage_2014; @blowes_geography_2019; @lyashevska_how_2012;
@hillebrand_biodiversity_2018], we expected that fish riverine communities
changes over time are highly variable and that the temporal trends of different
community metrics are only poorly related. Finally, given the high spatial
heterogeneity of river systems, we expected in turn a strong spatial
heterogeneity in community changes at small spatial scale.



# Results

```{r}
get_global_effect <- function (
  effect = inla_no_drivers_effects,
  resp = NULL,
  ci_lvl = "level:0.95"
  ) {

  out <- effect %>% 
    filter(
      ci_level == ci_lvl,
      term == "log1_year_nb",
      response == resp 
    )

  out[, c("low", "high", "mean")] %>%
    pivot_longer(everything()) %>%
    deframe()
}

p_ci <- function(x, r = 2, p = TRUE) {
  out <- round(x, r)

  if (p) {
    paste0(out["mean"], "%", " [", out["low"],"%,", out["high"],"%]")
  } else {
    paste0(out["mean"], " [", out["low"],",", out["high"],"]")
  }

}

get_effect_ci <- function(
  effect = tu,
  resp = NULL,
  term = "log1_year_nb",
  ci_lvl = "level:0.95",
  r = 2
  ) {
  term1 <- term

  out <- effect %>%
    filter(
      ci_level == ci_lvl,
      term == term1,
      response == resp
    )

  out <- out[, c("low", "high", "mean")] %>%
    pivot_longer(everything()) %>%
    deframe()

  p_ci(x = out, r = r, p = FALSE)

}
```

```{r}
trends <- map(unique(inla_no_drivers_effects$response),
  ~get_global_effect(
  effect = inla_no_drivers_effects,
  resp = .x,
  ci_lvl = "level:0.95"
  )
)
names(trends) <- unique(inla_no_drivers_effects$response)

trends80 <- map(unique(inla_no_drivers_effects$response),
  ~get_global_effect(
  effect = inla_no_drivers_effects,
  resp = .x,
  ci_lvl = "level:0.80"
  )
)
names(trends80) <- unique(inla_no_drivers_effects$response)

trends90 <- map(unique(inla_no_drivers_effects$response),
  ~get_global_effect(
  effect = inla_no_drivers_effects,
  resp = .x,
  ci_lvl = "level:0.90"
  )
)
names(trends90) <- unique(inla_no_drivers_effects$response)
```

```{r}
abun_trends <- p_ci(
  log_beta_to_perc_rate(trends$log_total_abundance)
  * log(10 + 1),
  r = 1)
rich_trends <- p_ci(
  log_beta_to_perc_rate(trends$log_chao_richness)
  * log(10 + 1),
  r = 1
)
hill_trends <- p_ci(
  trends$hillebrand_dis_scaled
  * log(10 + 1),
  r = 2, p = FALSE
)
exo_trends <- map_chr(c("perc_exo_sp", "perc_exo_abun"),
  ~p_ci(trends[[.x]] * log(10 + 1), r = 3, p = FALSE)
)
names(exo_trends) <- c("perc_exo_sp", "perc_exo_abun")

exo_trends90 <- map_chr(c("perc_exo_sp", "perc_exo_abun"),
  ~p_ci(trends90[[.x]] * log(10 + 1), r = 4, p = FALSE)
)
names(exo_trends90) <- c("perc_exo_sp", "perc_exo_abun")

exo_trends80 <- map_chr(c("perc_exo_sp", "perc_exo_abun"),
  ~p_ci(trends80[[.x]] * log(10 + 1), r = 3, p = FALSE)
)
names(exo_trends80) <- c("perc_exo_sp", "perc_exo_abun")
```

```{r}
extract_pred <- function (x = NULL, resp = NULL, val_to_extract = NULL) {

  ci_low <- x[[resp]]$quant0.025[val_to_extract]
  ci_high <- x[[resp]]$quant0.975[val_to_extract]

  # Does the ci at 95 overlap? 
  ci_overlap <- ci_low[val_to_extract[1]] > ci_high[val_to_extract[2]] |
    ci_high[val_to_extract[1]] < ci_low[val_to_extract[2]]

  list(
    val = x[[resp]]$mean[val_to_extract],
    ci_overlap = !ci_overlap
  )
}
#' Get percentage of difference, the first value being the reference
get_ratio <- function(x, r = 1) {
  #reduce(x, `/`)
  out <- round((x[2] - x[1]) / x[1] * 100, r)
  paste0(out, "%")
}
```

```{r}
tar_load(pred_number)
```

```{r}
pred_hft93_t0 <- map(
  setNames(names(pred_number$pred_hft93_t0), names(pred_number$pred_hft93_t0)),
  ~extract_pred(
    x = pred_number$pred_hft93_t0,
    resp = .x,
    val_to_extract = c("intact", "max")
  )
)

abun_rich_hft93_int_max <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft93_t0[[.x]]$val), r = 0)
)
other_hft93_int_max <- map_chr(
  setNames(names(pred_number$pred_hft93_t0), names(pred_number$pred_hft93_t0))[
    !names(pred_number$pred_hft93_t0) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft93_t0[[.x]]$val, r = 0)
  )
```


```{r}
pred_hft9309_t0 <- map(
  setNames(names(pred_number$pred_hft9309_t0), names(pred_number$pred_hft9309_t0)),
  ~extract_pred(
    x = pred_number$pred_hft9309_t0,
    resp = .x,
    val_to_extract = c("div/2", "inc/2")
  )
)

abun_rich_hft93909_dec_inc_by2 <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft9309_t0[[.x]]$val), r = 0)
)
other_hft93909_dec_inc_by2 <- map_chr(
  setNames(names(pred_number$pred_hft9309_t0), names(pred_number$pred_hft9309_t0))[
    !names(pred_number$pred_hft9309_t0) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft9309_t0[[.x]]$val, r = 0)
  )
```

```{r}
pred_hft93_t10 <- map(
  setNames(names(pred_number$pred_hft93_t10), names(pred_number$pred_hft93_t10)),
  ~extract_pred(
    x = pred_number$pred_hft93_t10,
    resp = .x,
    val_to_extract = c("intact", "max")
  )
)

abun_rich_hft93_int_max_t10 <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft93_t10[[.x]]$val), r = 0)
)
other_hft93_int_max_t10 <- map_chr(
  setNames(names(pred_number$pred_hft93_t10), names(pred_number$pred_hft93_t10))[
    !names(pred_number$pred_hft93_t10) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft93_t10[[.x]]$val, r = 0)
  )
```

```{r}
pred_hft9309_t10 <- map(
  setNames(names(pred_number$pred_hft9309_t10), names(pred_number$pred_hft9309_t10)),
  ~extract_pred(
    x = pred_number$pred_hft9309_t10,
    resp = .x,
    val_to_extract = c("div/2", "inc/2")
  )
)

abun_rich_hft93909_dec_inc_by2_t10 <- map_chr(
  setNames(c("log_total_abundance", "log_chao_richness"),
    c("log_total_abundance", "log_chao_richness")),
  ~ get_ratio(exp(pred_hft9309_t10[[.x]]$val), r = 0)
)
other_hft93909_dec_inc_by2_t10 <- map_chr(
  setNames(names(pred_number$pred_hft9309_t10), names(pred_number$pred_hft9309_t10))[
    !names(pred_number$pred_hft9309_t10) %in% c("log_chao_richness",
      "log_total_abundance")],
  ~ get_ratio(pred_hft9309_t10[[.x]]$val, r = 0)
  )
```

```{r}
tar_load(c(gaussian_inla_std_effects, gaussian_inla_exo_std_effects))
tu <- rbind(gaussian_inla_std_effects, gaussian_inla_exo_std_effects)
ci80 <- tu %>%
  filter(ci_level == "level:0.80")
ci90 <- tu %>%
  filter(ci_level == "level:0.90")
ci95 <- tu %>%
  filter(ci_level == "level:0.95")
```

- TODO:
  - explain higher spatial richness, past anthropogenic pressures: non native?
  - add stream gradient 

## Global temporal trends

Using a model without drivers but time, we found strong evidence for a global
increase in both total abundance and species richness of riverine fish
communities, respectively `r abun_trends` and `r rich_trends` by decade (CI
95%). We further found moderate evidence for a decline in the proportion of
non-native abundance, `r exo_trends[["perc_exo_abun"]]` by decade (CI 90%) but
no evidence for a net temporal trend in the proportion of non-native richness
(`r exo_trends[["perc_exo_sp"]]`, CI 80%). We also found that temporal
dissimilarity metrics of Jaccard, turnover and Simpson-based have overall effect
comparable in magnitude (Fig. S10), indicating the temporal changes of fish
communities are driven by both species replacement and changes in dominance
structure.

## Effect of anthropogenic pressures over space and time

We found strong evidence that past anthropogenic pressures, quantified by the
human footprint index at the year 1993, result in higher species richness (Fig.
1), as most degraded sites have `r abun_rich_hft93_int_max["log_chao_richness"]`
more richness than intact sites. Sites with higher past anthropogenic pressures
also had a higher proportion of non-native richness and non-native abundance
(strong evidence, Fig. 1), such as most degraded sites had a proportion of
non-native species twice higher than intact sites (resp.  
`r scales::percent(pred_hft93_t0$perc_exo_sp$val["max"])` and 
`r scales::percent(pred_hft93_t0$perc_exo_sp$val["intact"])`) and a proportion
of non-native abundance three times higher 
(`r scales::percent(pred_hft93_t0$perc_exo_abun$val["max"])` and
`r scales::percent(pred_hft93_t0$perc_exo_abun$val["intact"])` resp.). 
We found also strong evidence that past anthropogenic pressures resulted in
lower total abundance (Fig. 1), such as most degraded sites had 
`r abun_rich_hft93_int_max["log_total_abundance"]` lower total abundance than
intact sites.

```{r}
hft93abun <- get_effect_ci(resp = "log_total_abundance", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
hft93rich <- get_effect_ci(resp = "log_chao_richness", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
hft93exor <- get_effect_ci(resp = "perc_exo_sp", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c93")
hft93exoa <- get_effect_ci(resp = "perc_exo_abun", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c93")
hft93hill <- get_effect_ci(resp = "hillebrand_dis_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
hft93turn <- get_effect_ci(resp = "turnover_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c93")
```
We found strong evidence that past anthropogenic pressures increased the
temporal trends of total abundance, species richness (resp. Credible Interval 95%:$\beta\prime =$ 
`r hft93abun` and `r hft93rich`, Fig. 1). We also found weak evidence that past
anthropogenic pressures increased the proportion of non-native richness over
time (CI 80%:$\beta\prime =$ `r hft93exor`, Fig. 1), but no evidence that they
affect the proportion of non-native abundance (CI 80%:$\beta\prime =$
`r hft93exoa`, Fig. 1). We also found strong evidence that past anthropogenic
pressures increased temporal dissimilarity and turnover over time, i.e. respectively
indicating changes in species dominance and species replacement (resp. CI
95%:$\beta\prime =$ `r hft93hill` and `r hft93turn`).

In contrast, we found strong evidence that recent increases in anthropogenic pressures,
quantified by the ratio of the human footprint index between 2009 and 1993,
resulted in both species richness and total abundance (Fig. 1). 
A multiplication by two of human footprint between 1993 and 2009 decreased
species richness by `r abun_rich_hft93909_dec_inc_by2["log_chao_richness"]` 
and total abundance by `r abun_rich_hft93909_dec_inc_by2["log_total_abundance"]`
compared to a division by two of human footprint in the same period. We also
found strong evidence that recent increase in anthropogenic pressures increased
the proportion of non-native species and abundances (Fig. 1), respectively by five times
(`r scales::percent(pred_hft9309_t0$perc_exo_sp$val["div/2"])` vs
`r scales::percent(pred_hft9309_t0$perc_exo_sp$val["inc/2"])` for sites whose
anthropogenic pressures recently decrease and increase respectively) and by
eight times (`r scales::percent(pred_hft9309_t0$perc_exo_abun$val["div/2"])` vs
`r scales::percent(pred_hft9309_t0$perc_exo_abun$val["inc/2"])`).

```{r}
hft09abun <- get_effect_ci(resp = "log_total_abundance", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c9309_log2_ratio", r = 3)
hft09rich <- get_effect_ci(resp = "log_chao_richness", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c9309_log2_ratio", r = 3)
hft09exor <- get_effect_ci(resp = "perc_exo_sp", ci_lvl = "level:0.90", term = "log1_year_nb:hft_ix_c9309_log2_ratio", r = 2)
hft09exoa <- get_effect_ci(resp = "perc_exo_abun", ci_lvl = "level:0.80", term = "log1_year_nb:hft_ix_c9309_log2_ratio", )
hft09hill <- get_effect_ci(resp = "hillebrand_dis_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c9309_log2_ratio")
hft09turn <- get_effect_ci(resp = "turnover_scaled", ci_lvl = "level:0.95", term = "log1_year_nb:hft_ix_c9309_log2_ratio")
```

We did not found evidence for an effect of recent anthropogenic
pressures on the temporal trends of both total abundance and species richness
(resp. CI 80%: $\beta\prime =$ `r hft09abun` and `r hft09rich`, Fig. 1). However, we found
moderate evidence that recent increase in anthropogenic pressures increase
the proportion in non-native richness over time (CI 90%: $\beta\prime =$ `r hft09exor`,
Fig. 1) but not the proportion of non-native abundance (CI 80%: $\beta\prime =$
`r hft09exoa`, Fig. 1). We found also strong evidence that recent increase in
anthropogenic pressures resulted in stronger species turnover and dissimilarity
(resp. CI 95%: $\beta\prime =$ `r hft09turn` and `r hft09hill`, Fig. 1).

```{r}
hft2abun <- get_effect_ci(resp = "log_total_abundance", ci_lvl = "level:0.95",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio", r = 3)
hft2rich <- get_effect_ci(resp = "log_chao_richness", ci_lvl = "level:0.80",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio", r = 3)
hft2exor <- get_effect_ci(resp = "perc_exo_sp", ci_lvl = "level:0.90",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio", r = 3)
hft2exoa <- get_effect_ci(resp = "perc_exo_abun", ci_lvl = "level:0.80",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio")
hft2hill <- get_effect_ci(resp = "hillebrand_dis_scaled", ci_lvl = "level:0.95",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio")
hft2turn <- get_effect_ci(resp = "turnover_scaled", ci_lvl = "level:0.95",
  term = "log1_year_nb:hft_ix_c93:hft_ix_c9309_log2_ratio")
```

We also found strong evidence for antagonistic effect between past and recent
changes in anthropogenic pressures on the temporal trends of total abundance
(CI 95%: $\beta\prime =$ `r hft2abun`, Fig.1), meaning that recent anthropogenic decreased more
temporal trends of total abundance in sites that were previously more degraded.
We did not find evidence for such interaction for species richness
(CI 80%: $\beta\prime =$ `r hft2rich`, Fig. 1). We also found moderate evidence for a
synergistic effect between past and recent anthropogenic pressures on the
increase of the proportion of non-native species, but not of non-native
abundance (resp. CI 90%: $\beta\prime =$ `r hft2exor` and CI 80% `r hft2exoa`,
Fig.1). Finally, we found strong evidence for a synergistic between past and
recent anthropogenic pressures on the increase of temporal dissimilarity and
turnover over time (resp. CI 95%: $\beta\prime =$ `r hft2hill` and `r hft2turn`,
Fig.1).

## Typology of community temporal trends

We found that the Principal Component Analysis necessitated four principal
components over the four community metrics to explain more than 90% of variation
in temporal trends (69% and 87% with respectively two and three principal
components), one axis for each community metrics, meaning that there was a low
collinearity among the temporal trends of community metrics (Fig. 2). The six
cluster of community changes over time were related to PCA axis.  Around 40% of
the sites were attributed to a cluster related to a medium temporal increase in
species richness and total abundance, as well a medium temporal dissimilarity
and turnover (Fig. 2). The second cluster (16% of the sites) was related to
communities displaying a high turnover, and a relatively moderate increase in
species richness, total abundance and dissimilarity over time. The third, fourth
and fifth clusters (resp. 13%, 12% and 9% of the sites, Fig. 2) were
related to an increase in species richness over time, a 
decrease in total abundance and to a decrease in species richness over time
respectively. The last cluster (7% of the sites) was related to communities
displaying relatively low changes over time.

## Spatial distribution of community temporal trends 

We found that the spatial distribution of clusters of community temporal trends
were relatively homogeneous across realms and more heterogeneous across
hydrographic basins (Fig. 3). We further found that for
all community metrics, the random variance in the temporal trends, i.e. in the slope
coefficient of the time predictor, was higher for the site identity (nested in
basin) than for hydrographic basin (Table S14). Expect for total abundance and
species richness, the variance attributed to site was at least twice higher than
the one attributed to basin (Table S14).

# Discussion

### Global trends

The global trends of community metrics that we found are in accordance with
previous studies. We found an increase of freshwater fish total abundance of `r
abun_trends` by decade, that is in accordance with total abundance temporal
trends found for aquatic insects in a recent meta-analysis [+11.3% by decade,
@van_klink_meta-analysis_2020] and with the increase in freshwater insect
occupancy found in UK since 1994 [@outhwaite_agriculture_2022]. We also found
global increase in species richness which contrasts with the no net temporal
trends found in a previous meta-analysis on terrestrial and marine assemblages,
but in accordance with results found on terrestrial plants
[@dornelas_assemblage_2014], but in disagreement with no net richness trends
found in another meta-analysis on terrestrial plants [@vellend_global_2013]. We
further found that fish community temporal changes are mainly driven by
turnover, i.e. species replacement, in accordance with a previous meta-analysis
on marine, freshwater and terrestrial assemblages [@blowes_geography_2019]. We
further found that the same effect of temporal dissimilarity quantified by
the Jaccard index and the one quantified by Simpson index which are respectively
based on presence/absence and relative abundance
[@hillebrand_biodiversity_2018], which suggests that the changes in the riverine
fish communities over time are accompanied by large changes in dominance
structure.

### Effect of anthropogenic pressures on community changes 

We found that both past and recent changes in anthropogenic pressures, assessed
with the human footprint index, left a strong but contrasted spatial signature.
Most degraded sites from past anthropogenic pressures had surprisingly half more
species richness than intact sites, but a proportion twice and third times
higher of non-native species and abundance respectively. In contrast, sites that
experienced a doubling in anthropogenic pressures between 1993 and 2009 had a
baseline species richness 
`r abun_rich_hft93909_dec_inc_by2["log_chao_richness"]` lower than sites that
experienced a twofold division of anthropogenic pressures in the same period,
but had also a proportion five times and eight times more higher of non-native
species and abundance respectively. Higher proportion of non-native species
related to higher anthropogenic pressures are in accordance with studies showing
that higher density of human population, urban areas and roads, all present in
the human footprint index, was already shown to increase non-native richness by
creating more introduction opportunity [@anas_drivers_2021;
@leprieur_fish_2008]. Anthropogenic pressures can also improve the success of
non-native species establishment via rapid disturbance, hydrological alteration,
eutrophication, which native species may not cope with, or they may lose their
prior competitive advantage due to local adaptation over non-native
[@bernery_freshwater_2022; @byers_impact_2002]. The higher baseline species
richness found in most degraded sites might be related to the higher proportion
of non-native and in accordance with a study showing that in Nearctic and
Palearctic realms, non-native introductions is more important than native
extirpation [@villeger_homogenization_2011], which should result in higher
overall species richness. Both past and recent increase in anthropogenic
pressures resulted in a lower total abundance is in accordance with previous
results on the effects of land-use on the total abundance and species richness
of terrestrial vertebrates and insects [@newbold_global_2015;
@outhwaite_agriculture_2022].
 
Legacy of past anthropogenic pressures resulted in a larger increase over time
in species richness and total abundance, suggesting a increase from past
disturbances. The global temporal trends that we found in riverine fish
communities is really close from the one found in a recent meta-analysis in
freshwater insects at global scale, i.e. +13% by decade vs +11% by decade in
@van_klink_meta-analysis_2020. This result also corroborates the increasing
occupancy of freshwater insect assemblages found in UK since 1994
[@outhwaite_agriculture_2022], freshwater animal LPI (Living Planet Index) in
Netherlands since 1990 [@van_strien_modest_2016], while the median baseline of
the timeseries in our study was 1997. Such recoveries have been suggested to be
due to the adoption of many water quality policies in the European Union and the
United States since 1970s [@langford_long-term_2009;
@van_klink_meta-analysis_2020; @outhwaite_agriculture_2022]. Further in
agreement with the recovery hypothesis, 92% of the sites included in the present
study can be considered as degraded, i.e. had a human footprint index superior
to four in 1993 [@williams_change_2020]. Past anthropogenic pressures also
resulted in a larger increase of the proportion of non-native richness over
time, suggesting that the increase in species richness over time partly due to
the increasing share of non-native species in most degraded sites. However, we
found no effect of past anthropogenic pressures on the proportion of non-native
abundance, which might be explained by the importance of residential time after
introduction in the becoming of non-native species [@bernery_freshwater_2022].
We found that past anthropogenic pressures increase temporal dissimilarity,
mainly by species replacement, i.e. turnover, and changes in species dominance
structure. This result is in line with the increase in terms of total abundance,
species richness and proportion of non-native species, suggesting that it
underlines strong reshuffling of fish riverine  communities. Our results are
also in agreement with previous studies showing that the introduction of
non-native species is an important driver of temporal dissimilarity and turnover
in fish riverine communities [@toussaint_historical_2014;
@kuczynski_concomitant_2018].  Higher presence of non-native species have been
proposed to be the result that they are often ubiquitous generalists, enables
them to establish in degraded ecosystems that native species cannot cope with
[@toussaint_historical_2014] or they are excluded by over competition and over
predation [@byers_impact_2002].  Changes in dominance in ecological communities
have been reported to change species interactions, alter coexistence among
species and ultimately alter ecosystem functioning
[@hillebrand_consequences_2008], so it could be that larger changes in dominance
are related to ongoing adaptation in degraded environments, with a significant
contribution of non-native species [@schlaepfer_potential_2011]. All together,
our results suggest that fish riverine communities might not recover from water
quality improvement or migitation effort as previously suggested for freshwater
insects [@van_klink_meta-analysis_2020; @outhwaite_complex_2020]. Because we
found that proportion of non-native species richness increased over time in
degraded ecosystems, in addition to the fact that these degraded ecosystems
contains three times more non-native species in terms of abundance and two times
more in terms of richness, associated with stronger species replacement and
changes in dominance over time, it suggests instead an on-going adaptation to
degraded environments. Our results suggest that non-native species may play a
significant role in this adaptation [@toussaint_historical_2014;
@byers_impact_2002], but tolerant native species probably contribute too
[@scott_native_2001; @mccune_gains_2013].  It resonates with a study finding
that high spatial homogenization of fish riverine communities in Nearctic and
Palearctic realms, due to non-native introductions rather than native
extirpation [@villeger_homogenization_2011], and also with findings that
urbanization can lead to both taxonomic and functional homogenization in
freshwater macroinvertebrate assemblages [@liu_land_2022]. Metacommunity models
also predicts that selection and dispersal leads to an increase in local
richness and spatial homogenization [@mouquet_community_2003]. Finally, the
interpretation of on-going recomposition of riverine fish communities over time
in degraded ecosystems resonates with evidences that local species richness
might not decrease but global richness does because of on going spatial
homogenization [@primack_biodiversity_2018; @olden_defining_2006]. However, a more
definitive interpretation of those increases in species richness and total
abundance in aquatic systems would require a precise assessment of habitat
quality and of the tolerance of the species composing the communities of these
past degraded areas. Furthermore, those high level of community changes in
degraded ecosystems might be also driven by the increasing effect of climate
change, whose effects was shown to be more intense in degraded areas in riverine
fish communities [@comte_climate_2021] and in terrestrial insect assemblages
[@outhwaite_agriculture_2022].
 
In contrast, we found that the more a site was already degraded, the more a
recent addition of anthropogenic pressure decreases the temporal trends of
species richness and total abundance, highlighting that recent and past
anthropogenic pressures have an antagonistic effect on the temporal trends of
species richness and total abundance, and that recovery from past pressures was
not happening in the sites most degraded. We also found that a recent increase
in anthropogenic pressures increases in turn the proportion of non-native
species richness as found for past anthropogenic pressures. Moreover, we found a
synergistic effect of both anthropogenic pressures, meaning that the effect of
recent anthropogenic pressures on the increase of the proportion of non-native
species richness over time is stronger in the most degraded ecosystems.
However, we did not find an effect on the temporal trends of the proportion of
non-native abundance, as for past anthropogenic pressures, possibly because of
the time lag between species introductions and their establishment
[@bernery_freshwater_2022]. We also found that recent increase in anthropogenic
pressures increases temporal dissimilarity both in term of changes of dominance
structure and species replacement. We further found synergistic effect of past
and recent anthropogenic pressures, meaning that the effect of recent
anthropogenic pressures on community composition changes over time was stronger
in already degraded locations. All together, our results are in accordance with
the mounting body of literature highlighting the importance of the pressure
timing on the dynamic of ecological communities [@mccomb_editorial_2020;
@brooks_combined_2019; @orr_towards_2020; @chen_threshold_2020].

### Typology and spatial scale of community changes 

We found that the temporal trends of community changes lies in different
dimensions, i.e. changes in species richness, total abundance, turnover or
medium/low temporal changes. These results are in accordance with previous studies
showing that temporal dissimilarity is poorly related to temporal trends of
species richness [@dornelas_assemblage_2014; @blowes_geography_2019;
@hillebrand_biodiversity_2018].

The first cluster which included almost half of the sites present medium
changes, characterised by small increase in species richness and total abundance
over time. It is in line with the global increase of species richness and total
abundance over time that we found in our study, related to the legacy of past
anthropogenic pressures. Interestingly, the second cluster associated with high
turnover, i.e. high species replacement, regroups sites that also display a
slight increase in species richness and total abundance over time. However, the
third, fourth, and fifth clusters which are respectively related to the
strongest increase in species richness, the strongest decrease in total
abundance, and the strongest decrease in species richness over time, all of them
were not related to particularly high turnover. It is in accordance with
previous studies showing tedious links between temporal dissimilarity and
species richness trends [@dornelas_assemblage_2014; @blowes_geography_2019;
@hillebrand_biodiversity_2018]. Moreover, we found that the cluster related to
the strongest decrease in species richness over time regroups sites that do not
display changes in total abundance in average and that the cluster related to
strongest decrease in total abundance over time regroups sites that even display
a slight increase in species richness in average. It adds to the body of
literature highlighting the multidimensional nature of biodiversity
[@lyashevska_how_2012; @chase_embracing_2018] and thus suggest that community
temporal trends should be assessed with multiple metrics to better document in
which ways biodiversity changes over time.

We found that the proportion of clusters of community changes over time was
relatively homogeneous across realms, i.e. Nearctic, Palearctic, and
Australasia. They were more heterogeneous across basin within realms,
particularly in Palearctic probably because illustrated basins belonged to three
different countries separated by the English channel and the Baltic sea.
Overall, we found that the random variance in temporal trends attributed to site
(nested in basin) was twice higher than the one attributed to hydrographic
basin, suggesting that community changes in fish riverine communities are highly
heterogeneous at small spatial scale. Furthermore, we found that the random
variance related to the spatial structure contributed much more to the variance
of the community metrics than the fixed effects such as time, stream gradient
and anthropogenic pressures. This result is in accordance with what have been
found in a previous study about the spatial scale of biodiversity
[@chase_embracing_2018], rising probably from missing predictors such as the
distance from post glaciation colonization that explains largely baseline
composition and species richness for example or related to the grain size of the
predictors describing site environment that is of four kilometers. Further, this
result is in line with the finding that effect sizes of biodiversity responses
to ecological drivers decreases with the increase of spatial scale and so that
aggregation of small scale response to higher scales might lead to
overestimation of effect of large scale [@chase_embracing_2018].

## 5 Conclusion and perspectives

We uncovered the relationships between community changes and anthropogenic
pressures, confirming previous results and confirming a recovery of aquatic
organisms from the legacy of past anthropogenic pressures
[@van_klink_meta-analysis_2020; @outhwaite_complex_2020]. However, our
results show that this recovery in species richness and total abundance is
accompanied by an increase in the proportion of non-native species and of an
increase of temporal changes in dominance and species replacement. We further
showed that recent increase in anthropogenic pressure counteract the recovery
from the past, especially in the most degraded ecosystems. Our result also
suggests that anthropogenic pressures had both effects on spatial and temporal
domain, which completed contrasted results found in space and time
[@dornelas_assemblage_2014; @blowes_geography_2019;
@van_klink_meta-analysis_2020; @newbold_global_2015; @cardinale_is_2018]. By
investigating relationships among temporal trends of community metrics, we
showed that it might be necessary to assess several dimensions of biodiversity
changes to get a more precise understanding of the scenario of community
changes. Further promising investigations include the assessment of the
potentially interacting effects between anthropogenic pressures and climate
change [@comte_climate_2021; @outhwaite_agriculture_2022], as well as a complement
analysis of temporal trends at the population level to further investigate the
respective contribution of non-native and native species in the reshuffling of
fish riverine communities over space and time.

# References {-}

<div id="refs"></div>
