---
title: "Material & Methods"
author: "Danet, A, Giam, X, Olden, J, Comte, L"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2:
    toc: true
  word_document:
bibliography: "bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "graphics/knitr-",
  fig.retina = 1.618, # Control using dpi
  fig.width = 6,  # generated images
  fig.pos = "t",  # pdf mode
  fig.align = "center",
  dpi = if (knitr::is_latex_output()) 72 else 300,
  out.width = "100%",
  dev = "svg",
  dev.args = list(png = list(type = "cairo-png")),
  optipng = "-o1 -quiet"
)

library(here)
library(tidyverse)
library(magrittr)
library(cowplot)
library(targets)

if (Sys.info()["login"] == "alain")
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib")
  )
source(here::here("R", "misc.R"))
source(here::here("R", "variable_shortcut.R"))
```

```{r}
tar_load(c(site_desc_loc, measurement))
tar_load(c(modelling_data, filtered_dataset_modelling, analysis_dataset))
tar_load(filtered_op_protocol_modelling)
tar_load(c(measurement_exo, modelling_data_exo))
tar_load(add_dataset_ref)

measurement_exo <- measurement_exo %>%
  filter(op_id %in% unique(modelling_data_exo$op_id))
```



```{r}
nsitetot <- length(unique(site_desc_loc$siteid))
nmeas <- nrow(measurement)
nsp <- length(unique(measurement$species))
nop <- length(unique(measurement$op_id))
```

```{r}
site_year <- filtered_dataset_modelling$site_quanti %>%
  filter(variable == "year") %>%
  left_join(filtered_dataset_modelling$location, by = "siteid") %>%
  mutate(
    span = max - min + 1,
    completeness = n / (span)
    )

fnsite <- nrow(site_year)
fnmeas <- nrow(filtered_dataset_modelling$measurement)
fnsp <- length(unique(filtered_dataset_modelling$measurement$species))
fnop <- length(unique(filtered_dataset_modelling$measurement$op_id))

mspan <- median(site_year$span)
mcpl <- scales::percent(median(site_year$completeness))
mbl <- median(site_year$min)

eco <- table(filtered_dataset_modelling$location %>%
  mutate(ecoregion = str_replace(ecoregion, "Neartic", "Nearctic")) %>%
  pull(ecoregion)
) / nrow(filtered_dataset_modelling$location)

per_eco <- scales::percent(table_to_vec(eco), accuracy = 1)

pays <- table(filtered_dataset_modelling$location$country
) / nrow(filtered_dataset_modelling$location)
pays_vec <- table_to_vec(pays)

per_pays <- scales::percent(sort(pays_vec, decreasing = TRUE), accuracy = 1)
per_s4pays <- scales::percent(sum(sort(pays_vec, decreasing = TRUE)[1:4]))

# Add database
```


# Fish community time series

We used the RivFishTime database [@comte_rivfishtime_2021], a compilation of
more than 12,000 time series of riverine fish community abundances, by species,
mainly covering western and northern Europe, northern America, and southeastern
Australia. We completed the database with time series from Canada and United
States (Table S1), following the same criteria than RivFishTime for integration
[@comte_rivfishtime_2021].

We selected data to ensure that each time series had at least 5 years of data
over a 10 year period as well as a consistent sampling protocol through time. As
several sites had been sampled using different sampling methods (e.g.
electrofishing, seining) and/or over different periods of the year, we selected
for each site only the sampling events that were performed using the most
frequent protocol (i.e. the mode) and within 1.5 month of the most frequently
sampled month (i.e. 45 days before or after). When there were several sampling
events the same year, we selected the sampling that took place at the closest
date from the mode sampled date of the site. We further checked that the
reported unit of abundance was homogeneous for each time series.

The data selection resulted in `r fnsite` fish community time series, totalling
`r fnop` sampling events, `r fnmeas` species abundance records, and `r fnsp`
freshwater fish species. The median time span was of `r mspan`
years, the median baseline of the time series was `r mbl`, and the median
completeness of the time series was of `r mcpl` (see Fig. S2 for the complete distribution).
The sites were mostly located in Palearctic (`r per_eco["Palearctic"]`),
Nearctic (`r per_eco["Nearctic"]`) and Australasia (`r per_eco["Australasia"]`).
Four countries gathered `r per_s4pays` of the sites, namely Great Britain
(`r per_pays["GBR"]`), France (`r per_pays["FRA"]`), Sweden (`r per_pays["SWE"]`),
and the United States (`r per_pays["USA"]`, Table S4).

# Biodiversity facets

To evaluate temporal changes in freshwater fish communities, we assessed changes
in several biodiversity facets related to species composition, species richness 
and total abundances (Table S7).

## Dissimilarity metrics

As changes in community composition can be attributed either to changes of
dominance or of rare species, we described temporal dissimilarity with both the
complement of the Jaccard index (J) and the Simpson-based dissimilarity index
($H_d$) [@hillebrand_biodiversity_2018]. The former index is based on
presence/absence and is simply the sum of species gains and losses over the
total number of species across two samples (eq. \@ref(eq:jac)). The latter is
based on species relative abundances and their variation across two samples (eq.
\@ref(eq:hill)). The Jaccard index quantifies the extent of immigration /
extinction processes while the Simpson-based dissimilarity index quantifies the
extent of change in species dominance [@hillebrand_biodiversity_2018].

$$
J = \dfrac{S_{gain} + S_{loss}}{S_{tot}}
  (\#eq:jac)
$$ 

with $S_{gain}$, $S_{loss}$, $S_{tot}$ being the numbers of immigrant,
extirpated and total species respectively.

$$
\begin{aligned}
H_d &= 1 - H \\
H_d &= 1 - \dfrac{\sum_i (p_i - p^{\prime}_i)^2}{\sum_i p_i^2 + \sum_i p^{\prime2}_i - \sum_i p_i p^{\prime}_i}
(\#eq:hill)
\end{aligned}
$$ 

with $i$: species $i$, $p$: relative abundance and $\prime$: the focal community

We further partitioned the Jaccard dissimilarity index into two sets
of complementary indices. The first set was Appearance and Disappearance,
respectively the proportion of immigrant species ($S_{gain} / S_{tot}$) and
the proportion of extirpated species ($S_{loss} / S_{tot}$). The second set
was Turnover ($J_t$) and Nestedness ($J_n$), respectively
$J_t = (2 * min(S_{loss}, S_{gain})) / (S_{common} + (2 * min(S_{loss},
S_{gain})))$ and $J_n = 1 - J_t$ [@baselga_betapart_2012], $S_{common}$ being
the number of species present in both communities. High Turnover values indicate
that the changes in community composition result from species replacement,
whereas high values of Nestedness indicate species gains or losses from a nested
community, i.e. that a community is a subset of the other
[@baselga_temporal_2015]. We took the first year of sampling of a community as
the reference community.

## Total abundance


```{r}
effort_unit <- table(
  filtered_op_protocol_modelling$unitabundance,
  is.na(filtered_op_protocol_modelling$sampledlength_m)
) / nrow(filtered_op_protocol_modelling)

u_no_eff <- scales::percent(
  table_to_vec(
    effort_unit[, colnames(effort_unit) == "TRUE"]
  )
)
u_tot_na_eff <- effort_unit[, colnames(effort_unit) == "TRUE"] /
  (effort_unit[, colnames(effort_unit) == "TRUE"] +
    effort_unit[, colnames(effort_unit) == "FALSE"])
per_u_tot_na_eff <- scales::percent(u_tot_na_eff)

u_eff <- scales::percent(
  table_to_vec(
    effort_unit[, colnames(effort_unit) == "FALSE"]
  )
)
u_eff <- scales::percent(table_to_vec(effort_unit[, 1]), accuracy = 1)

unitvec <- table(filtered_op_protocol_modelling$unitabundance) /
  nrow(filtered_op_protocol_modelling)

u_per <- scales::percent(table_to_vec(unitvec))
```

Total abundance was reported in number of individuals (`r u_per["Count"]` of the sampling
events), density of individuals per 100 $m^2$ (`r u_per["Ind.100m2"]`), Catch Per
Unit Effort (`r u_per["CPUE"]`), and Leslie index (`r u_per["Leslie_index"]`,
Table S3).
Although we selected for strict protocol consistency, 70% or more of the
sampling events by unit of abundance did not reported sampling effort,
preventing us to harmonize count, abundance density and Catch Per Unit Effort
[@comte_rivfishtime_2021].

##  Species richness

```{r}
#see adjust_abun_chao()
cor_chao_rich <-  cor(
  modelling_data$chao_richness,
  modelling_data$species_nb,
  method = "spearman") %>%
round(., 2)

cor_log_chao_rich <-  cor(
  modelling_data$log_chao_richness,
  modelling_data$log_species_nb,
  method = "spearman") %>%
round(., 2)
```
As sampled species richness is a negatively biased estimator of the "true"
species richness, we corrected sampled species richness with the coverage-based
rarefaction and extrapolation methodology [@chao_coverage-based_2012]. The
estimated coverage of a sample is respectively positively and negatively related
to the number of individuals and number of singletons. For species richness to
be comparable across samples, we fixed the coverage of all samples at 98.5%.

As `r u_per["Ind.100m2"]` of the abundances were measured as density by 100
$m^2$ and `r u_per["CPUE"]` as Catch per Unit Effort, we did not have direct access
to the number of individuals and number of singletons to perform the
computation. In this case, we first divided each species abundance ($x_i$) by
the minimum values of abundance in the community (i.e. $x^{\prime}_i =
1 / min(x_i)$), which we further rounded so that each community had at least one singleton
species, i.e. a species with one individual. The correlation was very high
between raw species richness and chao richness (Spearman's $\rho = `r cor_chao_rich`$ for
both raw variables and log transformed ones, Fig. S3).

# Stream gradient and anthropogenic pressures

```{r}
n_na_riveratlas <- analysis_dataset[is.na(analysis_dataset$log_dist_up_km), ] %>%
  distinct(siteid) %$% nrow(.)
v_rvat <- scales::percent(
  (nrow(analysis_dataset) - n_na_riveratlas) /
    nrow(analysis_dataset)
  )

nb_op9309 <- sum(
  filtered_op_protocol_modelling$year >= 1993 &
  filtered_op_protocol_modelling$year <= 2009
)
perc_op9309 <- scales::percent(nb_op9309 / nrow(filtered_op_protocol_modelling))

nb_opb93 <- sum(
  filtered_op_protocol_modelling$year < 1993
)
perc_opb93 <- scales::percent(nb_opb93 / nrow(filtered_op_protocol_modelling))
nb_opa09 <- sum(
  filtered_op_protocol_modelling$year > 2009
)
perc_opa09 <- scales::percent(nb_opa09 / nrow(filtered_op_protocol_modelling))
```

```{r}
tar_load(pca_riv_str)
perc_var_rc1 <- pca_riv_str$rotated$Vaccounted["Proportion Var", "RC1"] %>%
  scales::percent(.)
tar_load(riveratlas_site)
env <- riveratlas_site[,
        colnames(riveratlas_site) %in%
          c("siteid", setNames(get_river_atlas_significant_var(), NULL))
        ] %>%
          sf::st_drop_geometry() %>%
          mutate(
              riv_str_rc1 =  pca_riv_str$rotated$scores[, "RC1"],
              riv_str_rc2 =  pca_riv_str$rotated$scores[, "RC2"]
          )
cor_env <- cor(env[, colnames(env) != "siteid"], method = "spearman") %>%
  round(., 2)
cor_env_rc1 <- cor_env["riv_str_rc1", ]
```


In dendritic networks, the environmental heterogeneity and connectivity along
the longitudinal (upstream-downstream) gradient strongly shape species
occurrences, immigration rates and community composition [ref].  To capture this
stream gradient, we described stream characteristics at each site by the
altitude (m), slope (deg.), average annual discharge ($m^3.s^-1$), distance from
source (km), and strahler order that we extracted from the HydroAtlas database
[@linke_global_2019; @robinson_earthenv-dem90_2014, Table S8]. We did so by
snapping the sites to the closest stream segment using a one kilometer buffer
(`r v_rvat` of the sites). We performed a Principal Component Analysis over the
site stream characteristics after log transforming (added absolute minimum
values plus one to avoid values $\leq 0$, $x^{\prime}_i = x_i + min(x_i) + 1$)
and standardizing all the variables, i.e. centering and scaling. We
orthogonally rotated the two first principal components, using the varimax
criterion [@kaiser_varimax_1958; @revelle_psych_2019], to increase the quality
of the variable representation (i.e. their loadings) on the two first principal
components. The first rotated component, related to average annual discharge,
distance from source and Strahler order and capturing `r perc_var_rc1` of the
variance (Fig. S4), was used as a composite variable describing the stream
gradient from upstream to downstream.

The degree of anthropogenic pressure was quantified using the human footprint index
[@venter_global_2016; @venter_sixteen_2016]. The human footprint index aggregates
an array of human pressures, including the extent of build environments, population density, 
electric infrastructures, croplands, pasture lands, roads, railways and
navigable pathways (Fig. S5). It does so by combining remote sensing data,
systematic surveys and modelling from ground data, making it less prone to errors
[@venter_global_2016]. The human footprint range between 0 and 50, with values
superior to four being considered in a degraded state [@williams_change_2020]. 
To capture both the effects of the legacy of past anthropogenic pressures
and its recent changes, we considered the human footprint index computed in 1993
and 2009 (i.e. 16 years span). Specifically, we used the human footprint index
of 1993 as a measure of the legacy of past anthropogenic pressures and the ratio
between the human footprint of 2009 and 1993 as a measure of the recent changes
in anthropogenic pressures. The human footprint values were extracted from the
HydroAtlas database at the reach scale [original resolution of 450meters,
@linke_global_2019]. In order to obtain interpretable coefficients of recent
changes in human footprint, we log-transformed the ratio of human footprint with
a base 2. Then, a value of minus one and one represent a division by two and a
multiplication by two of the human footprint between 1993 and 2009,
respectively. Only `r perc_opb93` of the samplings took place before 1993, while
`r perc_op9309` took place between 1993 and 2009 and `r perc_opa09` after 2009
(Fig. S1).
The human footprint indexes 1993 and 2009 were not correlated with the stream
gradient (Spearman's $\rho$, `r cor_env_rc1["hft_ix_c93"]` and
`r cor_env_rc1["hft_ix_c09"]` respectively).

```{r}
# Resolution of hydrorivers:
#https://www.hydrosheds.org/products/hydrorivers
```


## Non-native species data

```{r}
nat_ori <- table(measurement_exo$native_exotic_origin) / nrow(measurement_exo)
nat_per <- scales::percent(table_to_vec(nat_ori), accuracy = 0.1)
nat_per2 <- scales::percent(table_to_vec(nat_ori), accuracy = 0.01)
```

```{r}
sp_country_handmade <- measurement_exo %>%
  filter(native_exotic_origin == "handmade") %>%
  distinct(species, country)
```

The biogeographic origin of the fish species describing whether species were
native or introduced to a given drainage basin was retrieved using the global
database of @tedesco_global_2017 (`r nat_per["tedesco"]` of the species
occurrences). For the sites
falling outside of the river basins provided in @tedesco_global_2017, such as
for the sites located close to the shore, we used the closest basin in the
country (mean distance = xxx). For species not included in a given drainage
basin, we determined the origin of the species at the country scale using
Fishbase [@froese_fishbase_2021] (`r nat_per["autofishbase"]` of species
occurrences). Given the extent of the United States, we specifically completed
the global database with the Nonindigenous Aquatic Species (NAS) database
developed by the
U.S. Geological Survey ([https://nas.er.usgs.gov/](https://nas.er.usgs.gov/)), at the state scale
(`r nat_per2["usgs"]` of the species occurrences). Finally, we completed the
remaining species origins at the country scale, using national atlases and
Fishbase data in neighboring countries, such as for *Piaractus brachypomus* and
*Rutilus rutilus* in the United States (`r nat_per["handmade"]` of the species
occurrences, Table S6).

We then estimated the percentage of non-native species for each sampling events,
both for species richness and abundances (Table S7).

# Statistical analysis

## General statistical model

We modelled the temporal trends of the different biodiversity facets ($Y$) as
dependant of time ($\beta_0Time_t$, eq. \@ref(eq:gen)) measured as the number
of years since the beginning of the sampling at each site with $t_0 = 0$, the
stream gradient measured by the rotated PCA axis over stream characteristics,
the legacy of past anthropogenic pressures measured by the human footprint index
of 1993, and the recent changes in anthropogenic pressures measured by the ratio
between the human footprint index of 2009 and 1993. We included all the
predictors as main effects ($\sum_{k=1}\beta_kX_k$) to capture the differences
in biodiversity facets attributed to spatial effects of the ecological drivers.
We further included interactions between time and the ecological drivers
($\sum_{k=0, l\neq k}\beta_{kl}X_kX_l$) to test how stream gradient and
anthropogenic pressures affect the temporal trends in
biodiversity facets. Finally, we included the triple interactions between time
and the pairs of other ecological drivers ($\sum_{k=0, m \neq
n\neq k}\beta_{kmn}X_kX_mX_n$) to test for the presence synergistic
or antagonistic effects of the stream gradient and anthropogenic pressures on
the temporal trends in biodiversity facets.

Furthermore, the statistical model (eq. \@ref(eq:gen)) was adapted according of
the nature of the variables, namely total abundance and dissimilarity metrics.
For total abundance, we added the measurement unit of abundance as a categorical
variable both as a main effect and in interaction with time
[@van_klink_meta-analysis_2020]. The temporal trends in the main text and
supplementary are adjusted for raw count, i.e. the reference factor level in
modelling was raw count. Because dissimilarity metrics at each site was 0 at the
beginning of the time series, we fixed the intercept was fixed to zero. The
dissimilarity metrics were relative to the site and bounded between 0 and 1, we
did not expect average difference in dissimilarity which are not due to
differences in the temporal trends. Then, we removed the main effect of
ecological drivers ($\sum_{k=1}\beta_kX_k$ term from eq. \@ref(eq:gen)), but
kept their effects on the temporal trends.

We accounted for the spatial structure of the data by adding random effects to
the intercept ($\alpha$) and the slope of the temporal trends ($\beta_0$) on the
basin identity ($n$) and on the site identity, nested in basin ($i|n$). The
random effects and the error terms were modelled as a Normal distribution of
mean 0 and variance ($\sigma^2$).



$$
\begin{aligned}
    Y_{i|n, t} &= \alpha + \beta_0Time_t + \sum_{k=1}\beta_kX_k + \sum_{k=0, l\neq k}\beta_{kl}X_kX_l + \sum_{k=0, m \neq l\neq k}\beta_{klm}X_kX_lX_m + \epsilon_{i|n, t}
  (\#eq:gen)
\end{aligned}
$$ 


- $\alpha = \alpha_0 + a_n + a_{i|n}$
- $\beta_0 = \mu + b_n + b_{i|n}$
- $k, l, m \in [1,2,3]$: ecological drivers including stream gradient, legacy of past and recent changes in anthropogenic pressures
- $a_n, a_{i|n}, b_n, b_{i|n}, \epsilon_{i|n, t} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: hydrographic basin, $i$: site $i$, $t$: time $t$

- suggestion:
  - How model abundance and dissimilarity metrics?
  - Choice of response distribution?
  - Variable transformation

All the response variables were modelled with a Gaussian distribution following
previous studies modelling temporal trends of community composition, species
richness and total abundance at the global scale [@dornelas_assemblage_2014;
@blowes_geography_2019; @van_klink_meta-analysis_2020]. Other error structures might
be more appropriate to model response variables bounded between 0 and 1 and
representing ratio of discrete numbers such as the dissimilarity metrics and the
percentage of non-native species, doing so allows to obtain easily interpretable
coefficients across all biodiversity facets (e.g. temporal trends are not
interpretable as rates of change when modelled using a logit scale such as when
using a beta distribution). In addition, @blowes_geography_2019 previously found
that slope coefficients estimated with a gaussian error and a beta error had a
spearman correlation superior to 0.90 and give qualitatively similar results. We
therefore believe that this choice is not likely to alter our conclusions.

## Variable transformations

```{r}
tar_load(tab_waic)
perc_dec_waic <- scales::percent(mean((tab_waic[["Log (Year nb + 1)"]] - tab_waic[["Year nb"]]) / tab_waic[["Year nb"]]))
dec_waic <- round(mean((tab_waic[["Log (Year nb + 1)"]] - tab_waic[["Year nb"]]))) 
```

We log-transformed the number of years as $\log(year + 1)$ as it improved the
quality of the model fitting to the data, decreasing the WAIC [Watanabe-Akaike
information criterion, @gelman_understanding_2014; @vehtari_practical_2017] by 
`r dec_waic` in average (`r perc_dec_waic`) (Table S11). It
suggests the presence of non-linearity in the temporal trends, which is
particularly expected in the case of bounded variables such as the dissimilarity
metrics. Then the rate of change per decade of a given biodiversity facet
is found for a time value of $\log(10 + 1)$, i.e. `r round(log(10 + 1), 2)`
[Point to the model prediction plots with
double abscissa scale (log + year)]. We further log-transformed with a base 2
the recent changes in anthropogenic pressures quantified by the ratio
of human footprint index between 2009 and 1993, i.e.
$\log_2(HFT_{2009} / HFT_{1993})$. Then, a value of minus one and one
represent respectively a division by two and a multiplication by two of the
human footprint between 1993 and 2009. We log-transformed total abundance and
Chao species richness, then the temporal trends are expressed in percentage by
unit of time (See Table S9 for a summary of variable transformations).  

### Variable standardization

In order to compare the magnitude of the effects of time, stream gradient, and
anthropogenic pressures among biodiversity facets and to compare the magnitude
of the effects of the predictors, we scaled both biodiversity facets and the
predictors by their standard deviation prior to the model evaluation.

As our models contain interactions, the individuals slope coefficients can be
difficult to interpret without centering the predictors around ecological
relevant values [@gelman_scaling_2008]. As an example, the average temporal trends
estimated by $\beta_0$ in eq.\@ref(eq:gen) can only be interpreted when all the $X_k
= 0$. Without centering, it means that $\beta_0$ is interpretable when the
legacy of anthropogenic pressures, measured by human footprint at the year 1993,
and the stream gradient, measured by the values of the rotated PCA axis over
stream characteristics, are equal to 0.  Hence, we centered the human footprint
at the year 1993 and the coordinates along the PCA axis around their average
values. Recent changes in anthropogenic pressures, measured by the ratio between
the human footprint index of 2009 and 1993, was not centered as 0 values
indicate no recent changes in anthropogenic pressures. Time variable was not
centered either because then the main effects of the ecological drivers
($\sum_{k=1}\beta_k$) are then interpreted as a baseline effect, i.e. when time
is equal to 0.

### Model evaluation and confidence intervals 

The models were evaluated in a Bayesian framework using Integrated Nested
Laplacian Approximation (INLA), which approximates the posterior distribution of
the parameters and then do not rely on Markov chains and Monte Carlo
simulations, and then is a computationally efficient method to evaluate Bayesian
models [@rue_approximate_2009; @rue_bayesian_2017].  When estimating conjointly
the temporal trends at multiple location, a advantage of the Bayesian approach
is the estimation of credible intervals around the temporal trends estimated at
each location. A second advantage is that it allows the computation of credible
intervals, which translates in the probability that an unobserved parameter
falls in a given interval, to the difference with frequentist approach
[@greenland_statistical_2016].  We computed the credible intervals at 80%, 90%
and 95% using Highest Posterior Density method [@hyndman_computing_1996]. When 0
is outside the credible intervals of the coefficients at 80%, 90% and 95%, they
can respectively be interpreted as weak, moderate and strong evidence of an
effect [@van_klink_meta-analysis_2020; @mastrandrea_guidance_2010]. The model
evaluation was performed with the `INLA` R package [@rue_approximate_2009].

### Bayesian priors

INLA models were evaluated with defaults uninformative priors. The prior
distribution of fixed coefficients followed a flat zero centered normal
distribution ($\mathcal{N}(\mu,\, \sigma^2) = \mathcal{N}(0,\, 1000)$). The
prior distribution of the random effects and the gaussian error
($\epsilon_{it}$, eq. \@ref(eq:gen)) followed a log gamma distribution with
shape and inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \,
5.10^-5)$). We there had to back-transformed the estimated coefficients in order
to obtained the standard deviations attributed to the random effects and the
gaussian error ($\sigma = 1 / \sqrt{\tau}$). We checked that the slope
coefficients, random effects and the temporal trends by basin and site were the
same than with a implementation in frequentist. Then, we concluded that the
quality of parameter inference did not suffer from the uninformative priors.

### Model validity

```{r}

```

We checked the model validity visually by plotting the fitted versus the
observed values (Fig. S6). We visually inspected the PIT and CPO (respectively
Posterior Integral Transform and Conditional Predictive Ordinate) distribution
to assess both the overall quality of fitting, and the frequency of outliers. 
We also computed $R^2$ to assess the quality of the fit of the bayesian models.
We computed the variance associated of each predicted values from their
posterior distribution (`inla.rmarginal` in `INLA` R package), following @gelman_r-squared_2019
recommandations to take in account variability associated with priors ($\theta$). We then
computed marginal ($R^2_m$)and conditional ($R^2_c$) R-squared, respectively associated to the variance
explained by fixed effects and the one explained by both fixed and random
effects [@nakagawa_general_2013]. We only included
random effect on intercept in the $R^2$ computation, i.e. the basin effect
($a_n$) and the site effect (a_{i|n}) [@nakagawa_general_2013;
@lahuis_explained_2014]. As we computed the variance of predicted values for
each observations, we reported the median and the standard deviation of
R-squared [Table S12, @gelman_r-squared_2019].

$$
R^2_m = Var_{fit} / (Var_{fit} + Var_{res}) = \sigma^2(\hat{y_i}) /
(\sigma^2(\hat{y_i}) + \sigma^2(y_i - \hat{y_i}))
$$

$$
R^2_c = (Var_{fit} + (a_n)^2 + (a_{i|n})^2) / (Var_{fit}  + (a_n)^2 + (a_{i|n})^2 + Var_{res}))
$$

$$ with $y_i$ and $\hat{y_i}$ being respectively the observations and the
predicted values, $Var_{fit}$ and Var_{res} being respectively the variance of
predictive means and the variance of the residuals, @gelman_r-squared_2019.
$a_n$ and a_{i|n} are respectively the random standard deviation on the
intercept associated to the hydrographic basin and the site.

There was very multicollinearity in the model, as all Variance Inflation Factors
were around 1 (Table S10).

## Assessing the dimensionality of temporal community changes

We performed a PCA over the temporal trends of biodiversity facets at the site
level to find the linear combination of variables that explained the most
variance and separates linearly uncorrelated variables. In complement, we
performed a clustering analysis to group together sites that displayed similar
temporal community changes. The temporal trends for each biodiversity facet and
site were extracted by the Best Linear Unbiased Prediction method. Prior to the
analysis, we adjusted the temporal trends of total abundance for the unit of
measurement with the main effect coefficients and the one in interaction with
temporal trends. We did not include the temporal trends of the variables
describing composition of non-native species as the they are part of
biodiversity [@schlaepfer_non-native_2018] and site temporal trends in
percentage of non-native species displayed very little variation, then it was of
little interest to include them in the dimensionality analysis.

We performed the trimmed k-means clustering method [@fritz_tclust_2012], a
robust clustering method because it avoids the identification of spurious
clusters. The method consists of trimming the $\alpha$ most outlying data while
taking in account the multidimentional structure, the number of dimension being
the number of the biodiversity facets. To choose a relevant number of clusters,
we plotted the trimmed log-likelihood of the function as a function of the
proportion of the most outlying data trimmed ($alpha$) (Fig. S7). We thus
selected a partition of temporal community changes in six clusters with
$\alpha=5\%$. We did not constraint the algorithm for the relative size of shape
of the clusters, as we had no apriori expectation about them. The clustering
algorithm was run for a minimum number of one hundred iterations and up to 125.
To further control for the quality of each fish community changes assignment to
a given cluster, we discarded any fish community for which the second best
cluster assignment is 50% better, we did so by comparing the degree of
affiliation to the clusters [@fritz_tclust_2012]. The clustering was performed
using `tclust` R package.


# Reproducibility, ROBITT assessment and open science statement

The manuscript and the supplementary materials are written in Rmarkdown, i.e.
combining code and text, and are available on github. We further implement a
code pipeline using the `targets` R package to ensure that all the code, the
data, the figures, the manuscript and the results are up to date.

- todo: Robitt assessment

# References {-}
