---
title: "Material & Methods"
author: "Danet, A, Giam, X, Olden, J, Comte, L"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document:
    toc: true
bibliography: "bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 1.618, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "t",  # pdf mode
                      fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300,
                      out.width = "100%",
                      dev = "svg",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")

library(here)
library(tidyverse)
library(magrittr)
library(cowplot)
library(targets)

if (Sys.info()["login"] == "alain")
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib")
  )
```

```{r}
tar_load(c(site_desc_loc, measurement))
tar_load(c(modelling_data, filtered_dataset_modelling, analysis_dataset))
tar_load(filtered_op_protocol_modelling)
tar_load(c(measurement_exo, modelling_data_exo))

measurement_exo <- measurement_exo %>%
  filter(op_id %in% unique(modelling_data_exo$op_id))
```


# Dataset and data selection

```{r}
nsitetot <- length(unique(site_desc_loc$siteid))
nmeas <- nrow(measurement)
nsp <- length(unique(measurement$species))
nop <- length(unique(measurement$op_id))
```


The RivFishTime database [@comte_rivfishtime_2021] describes the time series of
riverine fish communities at large scale. It currently  holds `r nmeas` abundance measurement of
`nsp` freshwater fishes species over `nop` sampling events distributed over 
`r nsitetot` sites. The minimum time span of a timeseries in the dataset is 10
years.

We selected data to guarantee the quality and the internal homogeneity of the timeseries.
In a given site, we selected the sampling events that used the most frequently
used protocol (i.e. the mode) and the most used unity of abundances. We
selected the samplings that reported the year, but also at minimum the month or
the quarter of the year (i.e., January to march, April to June, etc.). For each site,  
we selected only the samplings that were performed within 1 month around the
most frequently sampled month, i.e. one month before or after. When there were
several samplings the same year, we selected the sampling that took place the
closest of the mode month/quarter of the site. Finally, we kept sites where the
number of samplings was equal or superior to five and that had time
span of at least 10 years.


# Characterise Biodiversity

Biodiversity was characterised by dissimilarity metrics, species richness and
total abundances.

## Dissimilarity metrics

As changes in community composition can be attributed either to abundant or rare
species, we described dissimilarity with both the complement of Jaccard (J) and
the Simpson-based dissimilarity index ($H_d$) [@hillebrand_biodiversity_2018].
The former index is based on presence/absence and is simply the sum of immigrant
and lost species over the total number of species across two samples (eq.
@eq:jac). The latter is based on species relative abundances and their variation
across two samples (eq. @eq:hill). Large trends for the Jaccard and
Simpson-based dissimilarity index indicates that changes in species composition
are mainly due to change in species dominance and identity
[@hillebrand_biodiversity_2018].

$$
J = \dfrac{S_{imm} + S_{lost}}{S_{tot}}
$$ {#eq:jac}

with $S_{imm}$, $S_{lost}$, $S_{tot}$ being the numbers of immigrant, lost and total species respectively.

$$
\begin{aligned}
H &= 1 - H \\
H &= 1 - \dfrac{\sum_i (p_i - p^{\prime}_i)^2}{\sum_i p_i^2 + \sum_i p^{\prime2}_i - \sum_i p_i p^{\prime}_i}
\end{aligned}
$$ {#eq:hill}

with $i$: species $i$, $p$: relative abundance and $\prime$: the focal community

In order to characterise the most precisely the changes in community
composition, we further partitioned the Jaccard dissimilarity index in two sets
of complementary indices. The first set was appearance and disappearance,
respectively the number of immigrant species ($\dfrac{S_{imm}}{S_{tot}}$) and
the number of disappearing species ($\dfrac{S_{lost}}{S_{tot}}$). The second set
was Turnover and Nestedness [@baselga_betapart_2012], respectively
($J_t = \dfrac{2 * min(S_{lost}, S_{imm})}{S_{common} + (2 * min(S_{lost}, S_{imm})}$) and ($J_n = 1 - J_t$).
High turnover values indicate that the changes in community composition are
hold by species replacement, while high values of Nestedness indicate species
gain or losses from a nested community, i.e. that a community is a subset of the
other [@baselga_temporal_2015].


## Total abundance

```{r}
table_to_vec <- function(x) {
  setNames(as.numeric(x), names(x))
}
```

```{r}
effort_unit <- table(
  filtered_op_protocol_modelling$unitabundance,
  is.na(filtered_op_protocol_modelling$sampledlength_m)
) / nrow(filtered_op_protocol_modelling)

u_no_eff <- scales::percent(
  table_to_vec(
    effort_unit[, colnames(effort_unit) == "TRUE"]
  )
)
u_tot_na_eff <- effort_unit[, colnames(effort_unit) == "TRUE"] /
  (effort_unit[, colnames(effort_unit) == "TRUE"] +
    effort_unit[, colnames(effort_unit) == "FALSE"])
per_u_tot_na_eff <- scales::percent(u_tot_na_eff)

u_eff <- scales::percent(
  table_to_vec(
    effort_unit[, colnames(effort_unit) == "FALSE"]
  )
)
u_eff <- scales::percent(table_to_vec(effort_unit[, 1]))

unitvec <- table(filtered_op_protocol_modelling$unitabundance) /
  nrow(filtered_op_protocol_modelling)

u_per <- scales::percent(table_to_vec(unitvec))
```

Total abundance was reported in raw count `r u_eff["Count"]`,
Abundance density per 100 $m^2$ (`r u_eff["Ind.100m2"]`),
Catch Per Unit Effort (`r u_eff["CPUE"]`), and Leslie index
(`r u_eff["Leslie_index"]`). Although
we selected for strict protocol consistency, 70% or more of the sampling
events by unit of abundance did not reported sampling effort, preventing us to
harmonize count, abundance density and Catch Per Unit Effort.


##  Species richness

```{r}
#see adjust_abun_chao()
cor_chao_rich <-  cor(
  modelling_data$chao_richness,
  modelling_data$species_nb,
  method = "spearman") %>%
round(., 2)

cor_log_chao_rich <-  cor(
  modelling_data$log_chao_richness,
  modelling_data$log_species_nb,
  method = "spearman") %>%
round(., 2)
```


As sampled species richness is a negatively biased estimator of the "true"
species richness, we corrected sampled species richness with the coverage-based
rarefaction and extrapolation [@chao_coverage-based_2012]. The coverage of a
sample depends both on the number of individuals and number of singletons. We
fixed the coverage of all samples to 98.5%.

Since `u_per["Ind.100m2"]` of the abundances were measured as density by 100
$m^2$ and `u_per["CPUE"]` by Catch per Unit of Effort, we did not direct have access
to the number of individuals and number of singletons. In this case, we divided
each species abundance ($x_i$) by the minimum values of abundance in the
community (i.e. $x^{\prime}_i = \dfrac{1}{min(x)}$) and then rounded the
obtained abundances. Then, those communities had at minimum one singleton
species. The correlation was very high raw species richness and chao richness
($\rho = `r cor_chao_rich`$ for both raw variables and log transformed ones).

# Environment and anthropogenic pressures

```{r}
n_na_riveratlas <- analysis_dataset[is.na(analysis_dataset$log_dist_up_km), ] %>%
  distinct(siteid) %$% nrow(.)
v_rvat <- scales::percent(
  (nrow(analysis_dataset) - n_na_riveratlas) /
    nrow(analysis_dataset)
  )

nb_op9309 <- sum(
  filtered_op_protocol_modelling$year >= 1993 &
  filtered_op_protocol_modelling$year <= 2009
)
perc_op9309 <- scales::percent(nb_op9309 / nrow(filtered_op_protocol_modelling))

nb_opb93 <- sum(
  filtered_op_protocol_modelling$year < 1993
)
perc_opb93 <- scales::percent(nb_opb93 / nrow(filtered_op_protocol_modelling))
nb_opa09 <- sum(
  filtered_op_protocol_modelling$year > 2009
)
perc_opa09 <- scales::percent(nb_opa09 / nrow(filtered_op_protocol_modelling))
```


Stream characteristics such as water flow, size of the streams, and
position in the dentritic network constitute tremendous constraints on fishes,
selecting the body shape of the fishes [ref], their life history traits [ref],
immigration rates [ref] of community and so ultimately community composition
[ref]. Such stream characteristics are often colinear and are often coined as 
"stream gradient". We described stream characteristics at each site by the
altitude (m), slope (deg.), average annual discharge ($m^3.s^-1$), distance from
source (km), and sthraler order that we extracted from 
the HydroAtlas database [@linke_global_2019, @robinson_earth-env_2014]. We did so by snapping the sites to
the closest stream segment at the condition that distance of the site to the closest
stream segment of the HydroAtlas database was inferior or equal to one kilometer
(`r v_rat` of the sites). We performed a PCA over the site stream
characteristics after log transformation (added absolute minimum values plus
one to avoid values $\leq 0$,
$x^{\prime}_i = x_i + min(x_i) + 1$) and standardization, i.e. centering and scaling
and log transformation of the variables. We rotated the two first principal
components with the varimax algorithm to improve the representativity of the
variables on the two first principal components.
The first rotated component was related to average annual discharge, distance
from source and sthraler order (Fig. S\@ref(fig:pca-riv-str)), the second
rotated component was related to stream altitude and slope. Then the first
rotated component was used as a composite variable characterising the stream
gradient from upstream to downstream.

Anthropogenic pressure was quantified with the Human footprint index
[@venter_global_2016, @venter_sixteen_2016]. The human footprint aggregates
human pressures namely as the extent of build environments, population density, 
electric infrastructure, croplands, pasture lands, roads, railways and
navigatable pathways. It does so by combining remote sensing data and
systematic surveys and modelling from the bottom up, making it less prone to errors
[@venter_global_2016]. The human footprint range between 0 and 50, values
exceding four being considered as a degraded state [@william_change_2020]. 
A study showed that the human footprint is related to the extinction risk of
mammals at world scale [@di_marco_changes_2018]. Two measures of human footprint
exists, in 1993 and in 2009 (i.e. 16 years span), enabling to characterise both
the legacy of Anthropogenic pressure and its recent changes. We took the human
footprint of 1993 as a measure of the legacy of anthropogenic pressure and the
ratio between the human footprint of 2009 and 1993. Only `r perc_opb93` of the
samplings took place before 1993, while `r perc_op9309` took place between 1993
and 2009 and `r perc_op9309` after 2009. The human footprint values were
extracted from the HydroAtlas database [@linke_global_2019].

The spatial resolution of human footprint and of the variables characterising
the stream gradient data was respectively 1 km and 3 arc degree. We extracted
HydroAtlas values at the reach level, which have a length of 4.2 km in average
[@linke_global_2019].

## Exotic species data

```{r}
nat_ori <- table(measurement_exo$native_exotic_origin) / nrow(measurement_exo)
nat_per <- scales::percent(table_to_vec(nat_ori), accuracy = 0.1)
nat_per2 <- scales::percent(table_to_vec(nat_ori), accuracy = 0.01)
```

```{r}
sp_country_handmade <- measurement_exo %>%
  filter(native_exotic_origin == "handmade") %>%
  distinct(species, country)
```



The origin status of the fish species were characterised
using a global database  at the basin scale [@tedesco_global_2017], for 
`r nat_per["tedesco"]` of the species occurrence. The sites that were not
perfectly included in a basin, as it was the case for sites near the shore, were 
matched to the closest basin in the country. For species missing of the
database, we determined the origin status of the species at the country scale
using fishbase [@froese_fishbase_2021] (`r nat_per["autofishbase"]` of species
occurrence.). Given the extent of the United States, we specifically completed
the global database with the Nonindigenous Aquatic Species (NAS) database of
USGS [https://nas.er.usgs.gov/](https://nas.er.usgs.gov/), at the state scale
(`r nat_per2["usgs"]` of the species occurrence). Finally, we completed the
remaining species status by hand (`r nat_per["handmade"]` of the species
occurrence) at the country scale. It was the case for occurrences of
*Piaractus brachypomus* and *Rutilus rutilus* in the United States for example.

We then characterised the composition in term of exotic species by the
percentage of species richness being exotic and the percentage of community 
abundance attributed to exotic species.

# Statistical analysis

## Temporal trends assessment

We modelled the temporal trends of biodiversity facets with hierarchical linear
models. All the response variables were modelled with a Gaussian distribution as
previous similar studies modelling temporal trends of community composition at
global scale [@dornalas_assemblage_2014; @blowes_geography_2019;
@klink_meta-analysis_2020]. Particularly the variables bounded between 0 and 1,
namely here dissimilarity metrics and percentages of exotic species,
@blowes_geography_2019 found that slope coefficients estimated with a gaussian error
and a beta error had a speaman correlation superior to 0.90 and give qualitately
similar results. Furthermore, the reponse variables are modelled on logit scale
(i.e. log(\dfrac{p}{1-p})) when assuming a beta error distribution and are
therefore interpretable at logit scale, i.e.  temporal trends are not
interpretable as rates of change [@blowes_geography_2019].

The value of each biodiversity facet ($Y$) at a site $i$ and time ($t$) was
modelled as a function of time ($\beta_0Time_t$), the covariates ($\sum_{k=1}\beta_kX_k$), the
interaction between time and covariates ($\sum_{k=0, n\neq k}\beta_{kn}X_kX_n$),
and the double interaction among time and the covariates ($\sum_{k=0, m \neq
n\neq k}\beta_{kmn}X_kX_mX_n$) plus an error term ($\epsilon_{it}$).
We added random effects on the intercept ($\alpha$) and the slope of the temporal
trends ($\beta_0$) such as they had random variations added to a main value
($\alpha_0$ and $\mu$ respectively). Random variations were added as dependant
on the basin identity ($n$) and on the site identity ($o$), nested in basin. The
random effects and the error terms were modelled as a Normal distribution of
mean 0 and variance ($\sigma$).


$$  
\begin{align}
    Y_{it} &= \alpha + \beta_0Time_t + \sum_{k=1}\beta_kX_k + \sum_{k=0, l\neq k}\beta_{kl}X_kX_l + \sum_{k=0, m \neq l\neq k}\beta_{klm}X_kX_lX_m + \epsilon_{it}
\end{align}
$$ {#eq:gen}


- $\alpha = \alpha_0 + a_n + a_o$
- $\beta_0 = \mu + b_n + b_o$
- $a_n, a_o, b_n, b_o, \epsilon_{it} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: basin, $o$: site within basin
- $i$: site $i$, $t$: time $t$

*Little cleaning to do in the notation: $Y_i|n,t and $i|n$ instead of $o$ ?*

The general model includes time and the ecological drivers, namely the PCA axis related to the stream gradient,
the Human footprint 1993, and the ratio of the Human footprint between 1993 and
2009 as main effects. We modelled the interactions among time and the ecological
drivers to test how the latter covariates affect the temporal trends of
biodiversity facets. Finally, we included the triple interactions among time and
the pairs of other covariates to test the presence synergistic or antagonistic
effects of the ecological drivers on the temporal trends of biodiversity facets.
 
In all the models, time was quantified as the number of years since the begining
of the sampling at each site with $t_0 = 0$. We log-transford the number of
years as $\Log(year + 1)$ to signicantly improved the quality of the fit of the
model on the data. We indeed found that log-transformed number of year increased   
the AIC by 29% in average from 0% for total abundance to 124% for the abundance
based dissimilarity index (Table S\@ref(tab:aic)) [missing chao richness and
percentage of exotic species in the comparison + refaire avec INLA?]. It
indicates the presense of non-linearity in the temporal trends, which is
particularly expected in the case of bounded variables such as the dissimilarity
metrics. In this situation, logging the predictor is a convenient way to linearise the
relationship. Then the rate of change of a given biodiversity facet for a decade
is found for a time log-transformed value of `r round(log(10 + 1), 2)`
approximatively (i.e. $\Log(10 + 1)$) [Point to the model prediction plots with
double abscisse scale (log + year)]. In order to obtain interpretable
coefficients of recent changes in human footprint, we Log-transformed the ratio
of human footprint with a base 2 (i.e.
$\Log_2(\dfrac{HFT_{2009}}{HFT_{1993}})$). Then, a value of minus one and one
represent respectively a division by two and a multiplication by two of the
human footprint between 1993 and 2009. 

### Statistical Models

[Add a table in supplementary with all the transformations of response and
explicative variables.]

We logged transformed total abundance and chao species richness in order to get
temporal trends expressed in percentage by unit of time
[@klink_meta-analysis_2020]. Furthermore, the general statistical model (eq.
@eq:gen) was adapted according of the nature of the variables, namely total
abundance and dissimilarity metrics. For total abundance, we added the
measurement unit of abundance as a categorical variable both as a main effect
and in interaction with time [@klink_meta-analysis_2020].  The temporal trends
in the main text and supplementary are presented for raw count, i.e. the
reference factor level in modelling was raw count.

The value of dissimilarity metrics at each site was 0 are the begining of the
temporal trends and the values were all bounded between 0 and 1. Then, we the
intercept was fixed to zero. Because the values of dissimilarity metrics are
relative to the site, we do not expect average difference in dissimilarity which
are not due to differences in the temporal trends. Then, we kept only the
ecological drivers as interactions with time,  i.e. we removed the
$\sum_{k=1}\beta_kX_k$ term from eq. @eq:gen.

### Variable standardization

In order to compare the magnitude of the temporal trends among biodiversity
facets and to compare the magnitude of the effects of the ecological drivers, we
scaled of the variables by their standard deviation prior to the model
evaluation. 

As our models contains interactions, the individuals slope coefficients can be
difficult to interpret without centering explicative variables around ecological
relevant values [@gelman_scaling_2008]. As an example, the average temporal trends
estimated by $\beta_0$ in eq. @eq:gen can only be interpreted when all the $X_k
= 0$. Without centering, it means that $\beta_0$ is interpretable when the
legacy of anthropogenic pressures, measured by human footprint at the year 1993,
and the values of PCA axis related to the stream gradient, are equal to 0.
Hence, we centered the human footprint at the year 1993 and the coordinates
along the PCA axis around their average values. Recent changes in anthropogenic
pressures, measured by the ratio between the two human footprint
($\Log_2(\dfrac{HFT_{2009}}{HFT_{1993}})$), was not centered as 0 values
indicate no recent changes in anthropogenic pressures. Time variable was not
centered either because then the main effects of the ecological drivers
($\sum_{k=1}\beta_k$) are then interpreted as a baseline effect, i.e. when time
is equal to 0.

### Model evaluation and confidence intervals 

The models were evaluated in a bayesian framework using Integrated Nested
Laplacian Approximation (INLA), which approximates the posterior distribution of
the parameters and then do not rely on Markov chains and Monte Carlo
simulations. We used default priors and checked that we obtain the same
estimation of slope coefficients, random effects and the same estimation of
temporal trends by basin and site (obtained by Best Linear Unbiased Prediction)
than with a implementation in frequentist with glmmTMB. When estimating
conjointly the temporal trends at multiple location, a avantage of the bayesian
approach over the frequentist one is that it allows the estimation of
confidence intervals around the temporal trends estimated at each location. A
second avantage is that it allows the computation of credible intervals at 80%,
90% and 95%. When 0 is outside the credible intervals of the coefficients at
80%, 90% and 95%, they can respectively be interpreted as weak, moderate and
strong evidence of an effect [@klink_meta-analysis_2020]. We computed
specifically with the credible intervals using the Highest Posterior Density
method, which fix the interval in the presence of asymetric posterior
distribution [ref].

### Model validity

TODO

### Sentivity of priors

comparison with glmmTMB, comparison with penalty complexity priors

## PCA and clustering of biodiversity temporal trends

TODO

### Pre-processing of temporal trends

adjustment of total abundance 

### Cluster identification

TODO

# Reproducibility, ROBITT assessment and open science statement


Robitt, Github, data repository, targets (proof of update), .Renv, Compendium

