---
title: "Material & Methods"
author: "Danet, A, Giam, X, Olden, J, Comte, L"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document:
    toc: true
bibliography: "bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 1.618, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "t",  # pdf mode
                      fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300,
                      out.width = "100%",
                      dev = "svg",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")

library(here)
library(tidyverse)
library(magrittr)
library(cowplot)
library(targets)

if (Sys.info()["login"] == "alain")
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib")
  )
```

```{r}
tar_load(c(site_desc_loc, measurement))
nsitetot <- length(unique(site_desc_loc$siteid))
nmeas <- nrow(measurement)
nsp <- length(unique(measurement$species))
nop <- length(unique(measurement$op_id))
```


# Dataset and data selection

The RivFishTime database [@comte_rivfishtime_2021] describes the time series of
riverine fish communities at large scale. It currently  holds `r nmeas` abundance measurement of
`nsp` freshwater fishes species over `nop` sampling events distributed over 
`r nsitetot` sites. The minimum time span of a timeseries in the dataset is 10
years.

We selected data to guarantee the quality and the internal homogeneity of the timeseries.
In a given site, we selected the sampling events that used the most frequently
used protocol (i.e. the mode) and the most used unity of abundances. We
selected the samplings that reported the year, but also at minimum the month or
the quarter of the year (i.e., January to march, April to June, etc.). For each site,  
we selected only the samplings that were performed within 1 month around the
most frequently sampled month, i.e. one month before or after. Finally, we kept sites where
the number of samplings was equal or superior to five and that had time
span of at least 10 years.


# Characterise Biodiversity

##  Species richness

## Total abundance

## Dissimilarity metrics

# Environment and anthropogenic pressures

# Statistical analysis

## Temporal trends assessment

### Statistical Models

### Standardization

### Model validity

### Sentivity of priors
comparison with glmmTMB, comparison with penalty complexity priors

## PCA and clustering of biodiversity temporal trends

### Pre-processing of temporal trends

adjustment 
### Cluster identification

# Reproducibility, ROBITT assessment and open science statement
Robitt, Github, data repository, targets (proof of update), .Renv, Compendium

