---
title: "Material & Methods"
author: "Danet, A, Giam, X, Olden, J, Comte, L"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  word_document:
  bookdown::pdf_document2:
    toc: true
bibliography: "bibliography.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "graphics/knitr-",
  fig.retina = 1.618, # Control using dpi
  fig.width = 6,  # generated images
  fig.pos = "t",  # pdf mode
  fig.align = "center",
  dpi = if (knitr::is_latex_output()) 72 else 300,
  out.width = "100%",
  dev = "svg",
  dev.args = list(png = list(type = "cairo-png")),
  optipng = "-o1 -quiet"
)

library(here)
library(tidyverse)
library(magrittr)
library(cowplot)
library(targets)

if (Sys.info()["login"] == "alain")
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib")
  )
source(here::here("R", "misc.R"))
```

```{r}
tar_load(c(site_desc_loc, measurement))
tar_load(c(modelling_data, filtered_dataset_modelling, analysis_dataset))
tar_load(filtered_op_protocol_modelling)
tar_load(c(measurement_exo, modelling_data_exo))

measurement_exo <- measurement_exo %>%
  filter(op_id %in% unique(modelling_data_exo$op_id))
```



```{r}
nsitetot <- length(unique(site_desc_loc$siteid))
nmeas <- nrow(measurement)
nsp <- length(unique(measurement$species))
nop <- length(unique(measurement$op_id))
```

```{r}
site_year <- filtered_dataset_modelling$site_quanti %>%
  filter(variable == "year") %>%
  left_join(filtered_dataset_modelling$location, by = "siteid") %>%
  mutate(
    span = max - min + 1,
    completeness = n / (span)
    )

fnsite <- nrow(site_year)
fnmeas <- nrow(filtered_dataset_modelling$measurement)
fnsp <- length(unique(filtered_dataset_modelling$measurement$species))
fnop <- length(unique(filtered_dataset_modelling$measurement$op_id))

mspan <- median(site_year$span)
mcpl <- scales::percent(median(site_year$completeness))
mbl <- median(site_year$min)

eco <- table(filtered_dataset_modelling$location %>%
  mutate(ecoregion = str_replace(ecoregion, "Neartic", "Nearctic")) %>%
  pull(ecoregion)
) / nrow(filtered_dataset_modelling$location)

per_eco <- scales::percent(table_to_vec(eco), accuracy = 1)

pays <- table(filtered_dataset_modelling$location$country
) / nrow(filtered_dataset_modelling$location)
pays_vec <- table_to_vec(pays)

per_pays <- scales::percent(sort(pays_vec, decreasing = TRUE), accuracy = 1)
per_s4pays <- scales::percent(sum(sort(pays_vec, decreasing = TRUE)[1:4]))
```


# Fish community time series

We used the RivFishTime database [@comte_rivfishtime_2021], a compilation of
more than 12,000 time series of riverine fish community abundances, by species,
mainly covering western and northern Europe, northern America, and southeastern
Australia. We completed the database with time series from Japan and United
States, following the same criteria than RivFishTime for integration
[@comte_rivfishtime_2021].

We selected data to ensure that each time series had at least 5 years of data
over a 10 year period as well as a consistent sampling protocol through time. As
several sites had been sampled using different sampling methods (e.g.
electrofishing, seining) and/or over different periods of the year, we selected
for each site only the sampling events that were performed using the most
frequent protocol (i.e. the mode) and within 1.5 month of the most frequently
sampled month (i.e. 45 days before or after). When there were several sampling
events the same year, we selected the sampling that took place at the closest
date from the mode sampled date of the site. We further checked that the
reported unit of abundance was homogeneous for each time series.

The data selection resulted in `r fnsite` fish community time series, totalling `r fnop` sampling
events, `r fnmeas` species abundance records, and `r fnsp`
freshwater fish species. The median time span was of `r mspan`
years, the median baseline of the time series was `r mbl`, and the median
completeness of the time series was of `r mcpl`. Finally, . The distribution of
those metrics are shown in supplementary figures (Fig. SX-X). The sites were
mostly located in Palearctic (`r per_eco["Palearctic"]`),
Nearctic (`r per_eco["Nearctic"]`) and Australasia (`r per_eco["Australasia"]`).
Four countries gathered `r per_s4pays` of the sites, namely Great Britain
(`r per_pays["GBR"]`), France (`r per_pays["FRA"]`), Sweden (`r per_pays["SWE"]`),
and the United States (`r per_pays["USA"]`).

# Biodiversity facets

To evaluate temporal changes in freshwater fish communities, we assessed changes
in several biodiversity facets related to species composition, species richness 
and total abundances.

## Dissimilarity metrics

As changes in community composition can be attributed either to changes of
dominance or of rare species, we described temporal dissimilarity with both the
complement of the Jaccard index (J) and the Simpson-based dissimilarity index
($H_d$) [@hillebrand_biodiversity_2018].  The former index is based on
presence/absence and is simply the sum of species gains and losses over the
total number of species across two samples (eq.  \@ref(eq:jac)). The latter is
based on species relative abundances and their variation across two samples (eq.
\@ref(eq:hill)). The Jaccard index quantifies the extent of immigration /
extinction processes while the Simpson-based dissimilarity index quantifies the
extent of change in species dominance [@hillebrand_biodiversity_2018].

$$
J = \dfrac{S_{gain} + S_{loss}}{S_{tot}}
  (\#eq:jac)
$$ 

with $S_{gain}$, $S_{loss}$, $S_{tot}$ being the numbers of immigrant,
extirpated and total species respectively.

$$
\begin{aligned}
H &= 1 - H \\
H &= 1 - \dfrac{\sum_i (p_i - p^{\prime}_i)^2}{\sum_i p_i^2 + \sum_i p^{\prime2}_i - \sum_i p_i p^{\prime}_i}
(\#eq:hill)
\end{aligned}
$$ 

with $i$: species $i$, $p$: relative abundance and $\prime$: the focal community

We further partitioned the Jaccard dissimilarity index into two sets
of complementary indices. The first set was Appearance and Disappearance,
respectively the proportion of immigrant species ($\dfrac{S_{gain}}{S_{tot}}$) and
the proportion of extirpated species ($\dfrac{S_{loss}}{S_{tot}}$). The second set
was Turnover ($J_t$) and Nestedness ($J_n$), respectively
($J_t = (2 * min(S_{loss}, S_{gain})) / (S_{common} + (2 * min(S_{loss},
S_{gain})))$) and ($J_n = 1 - J_t$) [@baselga_betapart_2012]. High Turnover
values indicate that the changes in community composition result from species
replacement, whereas high values of Nestedness indicate species gains or losses
from a nested community, i.e. that a community is a subset of the other
[@baselga_temporal_2015].

## Total abundance


```{r}
effort_unit <- table(
  filtered_op_protocol_modelling$unitabundance,
  is.na(filtered_op_protocol_modelling$sampledlength_m)
) / nrow(filtered_op_protocol_modelling)

u_no_eff <- scales::percent(
  table_to_vec(
    effort_unit[, colnames(effort_unit) == "TRUE"]
  )
)
u_tot_na_eff <- effort_unit[, colnames(effort_unit) == "TRUE"] /
  (effort_unit[, colnames(effort_unit) == "TRUE"] +
    effort_unit[, colnames(effort_unit) == "FALSE"])
per_u_tot_na_eff <- scales::percent(u_tot_na_eff)

u_eff <- scales::percent(
  table_to_vec(
    effort_unit[, colnames(effort_unit) == "FALSE"]
  )
)
u_eff <- scales::percent(table_to_vec(effort_unit[, 1]), accuracy = 1)

unitvec <- table(filtered_op_protocol_modelling$unitabundance) /
  nrow(filtered_op_protocol_modelling)

u_per <- scales::percent(table_to_vec(unitvec))
```

Total abundance was reported in number of individuals (`r u_per["Count"]` of the sampling
events), density of individuals per 100 $m^2$ (`r u_per["Ind.100m2"]`), Catch Per
Unit Effort (`r u_per["CPUE"]`), and Leslie index (`r u_per["Leslie_index"]`).
Although we selected for strict protocol consistency, 70% or more of the
sampling events by unit of abundance did not reported sampling effort,
preventing us to harmonize count, abundance density and Catch Per Unit Effort
[@comte_rivfishtime_2021].

##  Species richness

```{r}
#see adjust_abun_chao()
cor_chao_rich <-  cor(
  modelling_data$chao_richness,
  modelling_data$species_nb,
  method = "spearman") %>%
round(., 2)

cor_log_chao_rich <-  cor(
  modelling_data$log_chao_richness,
  modelling_data$log_species_nb,
  method = "spearman") %>%
round(., 2)
```


As sampled species richness is a negatively biased estimator of the "true"
species richness, we corrected sampled species richness with the coverage-based
rarefaction and extrapolation methodology [@chao_coverage-based_2012]. The
estimated coverage of a sample is respectively positively and negatively related
to the number of individuals and number of singletons. For species richness to
be comparable across samples, we fixed the coverage of all samples at 98.5%.
HERE

As `r u_per["Ind.100m2"]` of the abundances were measured as density by 100
$m^2$ and `r u_per["CPUE"]` by Catch per Unit of Effort, we did not direct have access
to the number of individuals and number of singletons. In this case, we divided
each species abundance ($x_i$) by the minimum values of abundance in the
community (i.e. $x^{\prime}_i = \dfrac{1}{min(x)}$) and then rounded the
obtained abundances. Then, those communities had at minimum one singleton
species. The correlation was very high raw species richness and chao richness
($\rho = `r cor_chao_rich`$ for both raw variables and log transformed ones).

# Environment and anthropogenic pressures

```{r}
n_na_riveratlas <- analysis_dataset[is.na(analysis_dataset$log_dist_up_km), ] %>%
  distinct(siteid) %$% nrow(.)
v_rvat <- scales::percent(
  (nrow(analysis_dataset) - n_na_riveratlas) /
    nrow(analysis_dataset)
  )

nb_op9309 <- sum(
  filtered_op_protocol_modelling$year >= 1993 &
  filtered_op_protocol_modelling$year <= 2009
)
perc_op9309 <- scales::percent(nb_op9309 / nrow(filtered_op_protocol_modelling))

nb_opb93 <- sum(
  filtered_op_protocol_modelling$year < 1993
)
perc_opb93 <- scales::percent(nb_opb93 / nrow(filtered_op_protocol_modelling))
nb_opa09 <- sum(
  filtered_op_protocol_modelling$year > 2009
)
perc_opa09 <- scales::percent(nb_opa09 / nrow(filtered_op_protocol_modelling))
```


Stream characteristics such as water flow, size of the streams, and
position in the dentritic network constitute tremendous constraints on fishes,
selecting the body shape of the fishes [ref], their life history traits [ref],
immigration rates [ref] of community and so ultimately community composition
[ref]. Such stream characteristics are often collinear and can be coined as 
"stream gradient". We described stream characteristics at each site by the
altitude (m), slope (deg.), average annual discharge ($m^3.s^-1$), distance from
source (km), and strahler order that we extracted from 
the HydroAtlas database [@linke_global_2019; @robinson_earthenv-dem90_2014]. We did
so by snapping the sites to the closest stream segment at the condition that
distance of the site to the closest
stream segment of the HydroAtlas database was inferior or equal to one kilometer
(`r v_rvat` of the sites). We performed a PCA over the site stream
characteristics after log transformation (added absolute minimum values plus
one to avoid values $\leq 0$,
$x^{\prime}_i = x_i + min(x_i) + 1$) and standardization, i.e. centering and scaling
and log transformation of the variables. We rotated the two first principal
components with the varimax algorithm to increase the quality of
variable representation on the two first principal components.
The first rotated component was related to average annual discharge, distance
from source and strahler order (Fig. S\@ref(fig:pca-riv-str)), the second
rotated component was related to stream altitude and slope. Then the first
rotated component was used as a composite variable characterising the stream
gradient from upstream to downstream.

Anthropogenic pressure was quantified with the Human footprint index
[@venter_global_2016; @venter_sixteen_2016]. The human footprint aggregates
human pressures namely as the extent of build environments, population density, 
electric infrastructure, croplands, pasture lands, roads, railways and
navigatable pathways. It does so by combining remote sensing data and
systematic surveys and modelling from the bottom up, making it less prone to errors
[@venter_global_2016]. The human footprint range between 0 and 50, values
exceding four being considered as a degraded state [@williams_change_2020]. 
A study showed that the human footprint is related to the extinction risk of
mammals at world scale [@di_marco_changes_2018]. Two measures of human footprint
exists, in 1993 and in 2009 (i.e. 16 years span), enabling to characterise both
the legacy of Anthropogenic pressure and its recent changes. We took the human
footprint of 1993 as a measure of the legacy of anthropogenic pressure and the
ratio between the human footprint of 2009 and 1993. Only `r perc_opb93` of the
samplings took place before 1993, while `r perc_op9309` took place between 1993
and 2009 and `r perc_op9309` after 2009. The human footprint values were
extracted from the HydroAtlas database [@linke_global_2019].

The spatial resolution of human footprint and of the variables characterising
the stream gradient data was respectively 1 km and 3 arc degree. We extracted
HydroAtlas values at the reach level, which have a length of 4.2 km in average
[@linke_global_2019].

## Exotic species data

```{r}
nat_ori <- table(measurement_exo$native_exotic_origin) / nrow(measurement_exo)
nat_per <- scales::percent(table_to_vec(nat_ori), accuracy = 0.1)
nat_per2 <- scales::percent(table_to_vec(nat_ori), accuracy = 0.01)
```

```{r}
sp_country_handmade <- measurement_exo %>%
  filter(native_exotic_origin == "handmade") %>%
  distinct(species, country)
```

The origin status of the fish species were characterised
using a global database  at the basin scale [@tedesco_global_2017], for 
`r nat_per["tedesco"]` of the species occurrence. The sites that were not
perfectly included in a basin, such as the sites located close to the shore, were 
matched to the closest basin in the country. For species missing of the
database, we determined the origin status of the species at the country scale
using fishbase [@froese_fishbase_2021] (`r nat_per["autofishbase"]` of species
occurrence.). Given the extent of the United States, we specifically completed
the global database with the Nonindigenous Aquatic Species (NAS) database of the
USGS [https://nas.er.usgs.gov/](https://nas.er.usgs.gov/), at the state scale
(`r nat_per2["usgs"]` of the species occurrence). Finally, we completed the
remaining species status by hand (`r nat_per["handmade"]` of the species
occurrence) at the country scale, such as *Piaractus brachypomus* and *Rutilus
rutilus* in the United States for example.

We then characterised the composition in term of exotic species by the
percentage of species richness being exotic and the percentage of community 
abundance attributed to exotic species.

# Statistical analysis

## Temporal trends assessment

We modelled the temporal trends of biodiversity facets with hierarchical linear
models. All the response variables were modelled with a Gaussian distribution as
previous similar studies modelling temporal trends of community composition,
species richness and total abundance at
global scale [@dornelas_assemblage_2014; @blowes_geography_2019;
@klink_meta-analysis_2020]. Particularly for the variables bounded between 0 and 1,
namely here dissimilarity metrics and percentages of exotic species,
@blowes_geography_2019 found that slope coefficients estimated with a gaussian error
and a beta error had a spearman correlation superior to 0.90 and give qualitatively
similar results. Furthermore, the response variables are modelled on logit scale
(i.e. $\log\dfrac{p}{1-p}$) when assuming a beta error distribution and are
therefore interpretable at logit scale, i.e.  temporal trends are not
interpretable as rates of change [@blowes_geography_2019].

The value of each biodiversity facet ($Y$) at a site $i$ and time ($t$) was
modelled as a function of time ($\beta_0Time_t$), the covariates ($\sum_{k=1}\beta_kX_k$), the
interaction between time and covariates ($\sum_{k=0, n\neq k}\beta_{kn}X_kX_n$),
and the double interaction among time and the covariates ($\sum_{k=0, m \neq
n\neq k}\beta_{kmn}X_kX_mX_n$) plus an error term ($\epsilon_{it}$).
We added random effects on the intercept ($\alpha$) and the slope of the temporal
trends ($\beta_0$) such as they had random variations added to a main value
($\alpha_0$ and $\mu$ respectively). Random variations were added as dependant
on the basin identity ($n$) and on the site identity ($o$), nested in basin. The
random effects and the error terms were modelled as a Normal distribution of
mean 0 and variance ($\sigma^2$).


$$
\begin{aligned}
    Y_{it} &= \alpha + \beta_0Time_t + \sum_{k=1}\beta_kX_k + \sum_{k=0, l\neq k}\beta_{kl}X_kX_l + \sum_{k=0, m \neq l\neq k}\beta_{klm}X_kX_lX_m + \epsilon_{it}
  (\#eq:gen)
\end{aligned}
$$ 


- $\alpha = \alpha_0 + a_n + a_o$
- $\beta_0 = \mu + b_n + b_o$
- $a_n, a_o, b_n, b_o, \epsilon_{it} \sim \mathcal{N}(0,\,\sigma^{2})$
- $n$: basin, $o$: site within basin
- $i$: site $i$, $t$: time $t$

*Little cleaning to do in the notation: $Y_i|n,t and $i|n$ instead of $o$ ?*

The general model includes time and the ecological drivers, namely the PCA axis related to the stream gradient,
the Human footprint 1993, and the ratio of the Human footprint between 1993 and
2009 as main effects. We modelled the interactions among time and the ecological
drivers to test how the latter covariates affect the temporal trends of
biodiversity facets. Finally, we included the triple interactions among time and
the pairs of other ecological drivers to test for the presence synergistic or antagonistic
effects of the ecological drivers on the temporal trends of biodiversity facets.
 
In all the models, time was quantified as the number of years since the begining
of the sampling at each site with $t_0 = 0$. We log-transford the number of
years as $\log(year + 1)$ to signicantly improved the quality of the fit of the
model on the data. We indeed found that log-transformed number of year increased   
the AIC by 29% in average from 0% for total abundance to 124% for the abundance
based dissimilarity index (Table S\@ref(tab:aic)) [missing chao richness and
percentage of exotic species in the comparison + refaire avec INLA?]. It
indicates the presense of non-linearity in the temporal trends, which is
particularly expected in the case of bounded variables such as the dissimilarity
metrics. In this situation, logging the predictor is a convenient way to linearise the
relationship. Then the rate of change of a given biodiversity facet for a decade
is found for a time log-transformed value of `r round(log(10 + 1), 2)`
approximatively (i.e. $\log(10 + 1)$) [Point to the model prediction plots with
double abscisse scale (log + year)]. In order to obtain interpretable
coefficients of recent changes in human footprint, we log-transformed the ratio
of human footprint with a base 2 (i.e.
$\log_2(\dfrac{HFT_{2009}}{HFT_{1993}})$). Then, a value of minus one and one
represent respectively a division by two and a multiplication by two of the
human footprint between 1993 and 2009. 

### Statistical Models

[Add a table in supplementary with all the transformations of response and
explicative variables.]

We logged transformed total abundance and chao species richness in order to get
temporal trends expressed in percentage by unit of time
[@klink_meta-analysis_2020]. Furthermore, the general statistical model (eq.
@ref(eq:gen)) was adapted according of the nature of the variables, namely total
abundance and dissimilarity metrics. For total abundance, we added the
measurement unit of abundance as a categorical variable both as a main effect
and in interaction with time [@klink_meta-analysis_2020].  The temporal trends
in the main text and supplementary are presented for raw count, i.e. the
reference factor level in modelling was raw count.

The value of dissimilarity metrics at each site was 0 are the begining of the
temporal trends and the values were all bounded between 0 and 1. Then, we fixed the
intercept was fixed to zero. Because the values of dissimilarity metrics are
relative to the site, we do not expect average difference in dissimilarity which
are not due to differences in the temporal trends. Then, we kept only the
ecological driver effects as interactions with time,  i.e. we removed the
$\sum_{k=1}\beta_kX_k$ term from eq.\@ref(eq:gen).

### Variable standardization

In order to compare the magnitude of the temporal trends among biodiversity
facets and to compare the magnitude of the effects of the ecological drivers, we
scaled of the variables by their standard deviation prior to the model
evaluation. 

As our models contains interactions, the individuals slope coefficients can be
difficult to interpret without centering explicative variables around ecological
relevant values [@gelman_scaling_2008]. As an example, the average temporal trends
estimated by $\beta_0$ in eq.\@ref(eq:gen) can only be interpreted when all the $X_k
= 0$. Without centering, it means that $\beta_0$ is interpretable when the
legacy of anthropogenic pressures, measured by human footprint at the year 1993,
and the values of PCA axis related to the stream gradient, are equal to 0.
Hence, we centered the human footprint at the year 1993 and the coordinates
along the PCA axis around their average values. Recent changes in anthropogenic
pressures, measured by the ratio between the two human footprint
($\log_2(\dfrac{HFT_{2009}}{HFT_{1993}})$), was not centered as 0 values
indicate no recent changes in anthropogenic pressures. Time variable was not
centered either because then the main effects of the ecological drivers
($\sum_{k=1}\beta_k$) are then interpreted as a baseline effect, i.e. when time
is equal to 0.

### Model evaluation and confidence intervals 

The models were evaluated in a bayesian framework using Integrated Nested
Laplacian Approximation (INLA), which approximates the posterior distribution of
the parameters and then do not rely on Markov chains and Monte Carlo
simulations. We used default priors and checked that we obtain the same
estimation of slope coefficients, random effects and the same estimation of
temporal trends by basin and site (obtained by Best Linear Unbiased Prediction)
than with a implementation in frequentist with glmmTMB. When estimating
conjointly the temporal trends at multiple location, a advantage of the bayesian
approach over the frequentist one is that it allows for the estimation of
confidence intervals around the temporal trends estimated at each location. A
second advantage is that it allows the computation of credible intervals at 80%,
90% and 95%. When 0 is outside the credible intervals of the coefficients at
80%, 90% and 95%, they can respectively be interpreted as weak, moderate and
strong evidence of an effect [@klink_meta-analysis_2020]. We specifically
computed the credible intervals using the Highest Posterior Density
method, which fix the interval in the presence of asymmetric posterior
distribution of the parameters [ref].

### Priors

INLA models were evaluated with defaults uninformative priors. The prior
distribution of fixed coefficients followed a flat zero centered normal
distribution ($\mathcal{N}(\mu,\, \sigma^2) = \mathcal{N}(0,\, 1000)$). The
prior distribution of the random effects and the gaussian error
($\epsilon_{it}$, eq. \@ref(eq:gen)) followed a log gamma distribution with shape and
inverse scale parameters ($\mathcal{G}(s, \, \tau) = \mathcal{G}(1, \, 5.10^-5)$). We there had to
back-transformed the estimated coefficients in order to obtained the standard
deviations attributed to the random effects and the gaussian error ($\sigma =
\dfrac{1}{\sqrt{\tau}}$). As the estimation of the main effects and the random
effects are of the same with INLA than in frequentist in glmmTMB [show the table!], we
concluded that the quality of parameter inference did not suffer from the
informative priors.

### Model validity

We checked the model validity visualy by plotting the fitted values versus the
observed [ref plot].

- Add plots of posterior marginal distribution of the parameters? (it would be like
  8*10 plots for the fixed effects only)
- Add PIT and CPO [@klink_meta-analysis_2020]

## PCA and clustering of the temporal trends of biodiversity facets

We performed a PCA over the temporal trends of biodiversity facets at the site
level in order to characterise the dimentionality of community changes. In
complement of this analysis, we performed a cluster analysis to isolate and
illustrate different groups of community changes. The temporal trends of biodiversity
facets at site scale were extracted by the Best Linear Unbiased Prediction
method. Prior to the analysis, we adjusted the temporal trends of total
abundance for the unit of measurement with the main effect coefficients and the
one in interaction with temporal trends. Furthermore, we did not include the
temporal trends of the variables describing composition of exotic species.

We performed the clustering using k-means and data trimming
[@fritz_tclust_2012], this method is part of the family of robust clustering
methods. The method consists on trimming the $\alpha$ most outlying data while
taking in account is multidimentional structure. To choose a relevant number of
clusters, we plotted the trimmed log-likelihood of the function in function of
the proportion of the most outlying data trimmed [ref plot]. We finally selected six  
clusters. We did not contraints the algorithm for the size the clusters, as we
had no apriori expectation about them and we used a restricting factor. We
started with 100 iterations and set the maximum to 125. To further control for
the quality of the site assignment on a cluster, we set cluster assignment as NA
if another cluster assignment would have been more 50% better. The clustering
was performed using `tclust` R package.


# Reproducibility, ROBITT assessment and open science statement

All the code to reproduce the analysis from the raw data to the manuscript is
available on github. We further implement a pipeline using the `targets` R
package to ensure that all the code, the data, and the results are up to date.

- todo: Robitt assessment

# References {-}
