---
title: "Supplementary Figures and Tables"
author: "Danet, A, Giam, X, Olden, J, Comte, L"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  bookdown::pdf_document2:
    df_print: paged
    toc: true
bibliography: "bibliography.bib"
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
---

```{r options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 1.618, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "t",  # pdf mode
                      fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300,
                      out.width = "100%",
                      dev = "svg",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")
```

```{r library, include = FALSE}
source(here::here("start_rmd.R"))
theme_set(theme_cowplot())
library(gt)
library(gtsummary)
library(ggbreak)
```

```{r}
get_tbl_summary <- function(df = NULL) {

t1 <- df %>%
  tbl_summary(statistic = all_continuous() ~ "{N_nonmiss}", missing = "no") %>%
  modify_header(stat_0 ~ "**N (no missing)**")

t2 <- df %>%
  tbl_summary(statistic = all_continuous() ~
    "{median} ({p25}, {p75})",
  missing = "no") %>%
  modify_header(stat_0 ~ "Median (1st Q, 3rd Q)")

t3 <- df %>%
  tbl_summary(statistic = all_continuous() ~
    "({min}, {max})",
  missing = "no") %>%
  modify_header(stat_0 ~ "(Min, Max)")

tbl_merge(list(t1, t2, t3)) %>%
  modify_footnote(everything() ~ NA_character_) %>%
  modify_spanning_header(everything() ~ NA_character_)

}
ecdf_fun <- function(x, perc) ecdf(x)(perc)
```

- Add plot species richness / chao richness

# Data selection


```{r}
tar_load(filtered_op_protocol_modelling)
tar_load(filtered_dataset_modelling)
tar_load(c(modelling_data, modelling_data_exo))
tar_load(c(clust_var, exo_resp_var))
tar_load(riveratlas_site)
tar_load(c(site_no_drivers_inla_tot, site_env, gaussian_inla_std_effects, gaussian_inla_exo_std_effects))

mykable <- function(data) {
  knitr::kable(data, booktabs = TRUE, digits = 2)
}
str_to_title
```


```{r tab-pro-ch}
tabyl_df(filtered_dataset_modelling$site_quali, "protocol") %>%
  mutate(
    protocol = str_replace_all(protocol, "_", " "),
    # For rotenone:
    protocol = str_replace_all(protocol, "([a-z])([A-Z])", "\\1 \\2")
    ) %>%
  rename_with(str_to_sentence) %>%
  kbl(booktabs = T,
    label = "tab-pro",
    caption = "Protocol occurrence across sites"
    ) %>%
  kable_styling(latex_options = "striped")
```


```{r tab-u-ch, fig.cap="Abundance unit occurrence in sites"}
tabyl_df(filtered_dataset_modelling$site_quali, "unitabundance") %>%
  mutate(
    unitabundance = str_replace_all(unitabundance, "_", " "),
    unitabundance = str_replace_all(unitabundance, get_unitabun_replacement())
    ) %>%
  rename_with(str_to_sentence) %>%
  rename(`Abundance unit` = Unitabundance) %>%
  kbl(booktabs = T,
    label = "tab-u",
    caption = "Abundance unit across sites"
    ) %>%
  kable_styling(latex_options = "striped")
```


```{r tab-loc-u, fig.cap=""}
location_site <- filtered_dataset_modelling$location %>%
  mutate(ecoregion = str_replace(ecoregion, "Neartic", "Nearctic")) %>%
  group_by(ecoregion, country)

location_summary <- location_site %>%
    summarise(n = n(), .groups = "drop") %>%
    mutate(
      percent = scales::percent(n / sum(n), accuracy = .01),
    ) %>%
    arrange(desc(ecoregion), desc(n)) %>%
    rename(realm = ecoregion) %>%
    rename_with(str_to_sentence)
location_summary %>%
  kbl(booktabs = T,
    label = "tab-loc",
    caption = "Realm and country occurrence across sites"
    ) %>%
  kable_styling(latex_options = "striped")
```

```{r}
year_op_cap <- "Distribution of year of samplings. Years 1993 and 2009 are
highlighted because they correspond to the two human footprint measurements."
```

```{r y-op, fig.cap=year_op_cap}
nb_op9309 <- sum(
  filtered_op_protocol_modelling$year >= 1993 &
  filtered_op_protocol_modelling$year <= 2009
)
perc_op9309 <- scales::percent(nb_op9309 / nrow(filtered_op_protocol_modelling))

filtered_op_protocol_modelling %>%
  ggplot(aes(x = year)) +
  geom_histogram() +
  geom_vline(
    xintercept = c(1993, 2009),
    lwd = 2, color = "red"
    ) +
  labs(x = "Year", y = "Count") +
  annotate("text", x = 2001, y = 8300,
    label = paste0(
      nb_op9309, " samplings between 1993 and 2009 (", perc_op9309,")"
      )
  )
```

```{r}
site_year <- filtered_dataset_modelling$site_quanti %>%
  filter(variable == "year") %>%
  left_join(filtered_dataset_modelling$location, by = "siteid") %>%
  mutate(
    span = max - min + 1,
    completeness = n / (span),
    cat_span = cut(span, c(0, 10, 20, 30, 40, 70), right = TRUE,
      include.lowest = TRUE
    ),
    cat_completeness = cut(
      completeness, c(0, .1, .25, .5, .75, 1), right = TRUE,
      include.lowest = TRUE
    )
    )
```

```{r}
p_baseline_year <- site_year %>%
  ggplot(aes(x = min)) +
  geom_histogram() +
  labs(x = "Baseline year", y = "Number of sites") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

p_span <- site_year %>%
  ggplot(aes(x = cat_span, fill = ecoregion)) +
  geom_bar() +
  labs(x = "Year span (year)", y = "Number of sites") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

p_completness <- site_year %>%
  ggplot(aes(x = cat_completeness, fill = ecoregion)) +
  geom_bar() +
  labs(
    x = "Completeness of sampling (nb sampling / year span )",
    y = "Number of sites"
  )
```

```{r}
cap <- "Distribution of (A) Baseline year, (B) year span and (3) completeness of
sampling by site."
```

```{r base-span-cpl, fig.cap=cap, fig.height = 12}
leg_ecoregion <- get_legend(p_span + theme(legend.position = "bottom"))
plot_grid(
  p_baseline_year,
  p_span + theme(legend.position = "none"),
  p_completness + theme(legend.position = "none"),
  leg_ecoregion,
  ncol = 1,
  label = "AUTO"
)
```


# Biodiversity summary metrics 

## Community metrics 

```{r tab-resp-ck}
tmp <- modelling_data %>%
  left_join(
    modelling_data_exo %>%
      select(siteid, year, perc_exo_sp, perc_exo_abun),
    by = c("siteid", "year")
  ) %>%
  select(all_of(c("chao_richness", "species_nb",
        clust_var, exo_resp_var)))
colnames(tmp) <- c(get_var_replacement(), get_model_term_replacement())[colnames(tmp)]
resp_summary <- tmp %>%
  pivot_longer(everything(),
    names_to = "Community metric",
    values_to = "values") %>%
  group_by(`Community metric`) %>%
  summarise(summ = list(enframe(summary_distribution(values)))) %>%
  unnest(summ) %>%
  pivot_wider(names_from = "name", values_from = "value") %>%
  mutate(
    median = round(median, 2),
    `Median (Q1, Q3)` = paste0(median, " (", round(`1st_quart`, 2), ",", round(`2nd_quart`, 2),
      ")"),
    `(Min, Max)` = paste0("(", round(min, 2), ",", round(max, 2), ")")
  ) %>%
  select(`Community metric`, n, `Median (Q1, Q3)`, `(Min, Max)`, median) %>%
  arrange(median) %>%
  select(-median)
  
resp_summary %>%
  kbl(booktabs = T,
    label = "tab-resp",
    caption = "Summary descriptors of response variables distribution"
    ) %>%
  kable_styling(latex_options = "striped")
```

## Exotic species

```{r}
tar_load(c(measurement_exo))
# filter
measurement_exo <- measurement_exo %>%
  filter(op_id %in% unique(modelling_data_exo$op_id))
```


```{r, tab_exo}
exo_sp_var <- c(
  "species_nb_nat", "species_nb_exo",
  "perc_exo_sp", "perc_nat_sp",
  "perc_exo_abun", "perc_nat_abun"
)

exotic_source_summary <-  measurement_exo[, "native_exotic_origin"] %>%
  group_by(native_exotic_origin) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(
    percent = scales::percent(n / sum(n), accuracy = .01),
    native_exotic_origin = get_nat_origin_replacement()[native_exotic_origin]
    ) %>%
  arrange(desc(n)) %>%
  rename(`Source for native species status` = native_exotic_origin) %>%
  rename_with(str_to_sentence)


tmp <- modelling_data_exo %>%
  select(all_of(exo_sp_var))
colnames(tmp) <- get_var_replacement()[colnames(tmp)]
exotic_compo_summary <- tmp %>%
  pivot_longer(everything(),
    names_to = "Community metric",
    values_to = "values") %>%
  group_by(`Community metric`) %>%
  summarise(summ = list(enframe(summary_distribution(values)))) %>%
  unnest(summ) %>%
  pivot_wider(names_from = "name", values_from = "value") %>%
  mutate(
    median = round(median, 2),
    `Median (Q1, Q3)` = paste0(median, " (", round(`1st_quart`, 2), ",", round(`2nd_quart`, 2),
      ")"),
    `(Min, Max)` = paste0("(", round(min, 2), ",", round(max, 2), ")")
  ) %>%
  select(`Community metric`, N = n, `Median (Q1, Q3)`, `(Min, Max)`) %>%
  arrange(`Community metric`)
```

```{r tab-exo-source-ck}
exotic_source_summary %>%
  kbl(booktabs = T,
    label = "tab-exo-source",
    caption = "Data source for native species status in term of species occurrence") %>%
  kable_styling(latex_options = "striped")
```

```{r tab-exo-ck}
exotic_compo_summary %>%
  kbl(booktabs = T,
    label = "tab-exo",
    caption = "Summary distribution of native and non-native species metrics") %>%
  kable_styling(latex_options = "striped")
```

# Environmental variables

```{r}
tar_load(pca_riv_str)
```


```{r, tab-evt, fig.cap="Descriptive statistics for the environmental variables"}
col_pca_riv <- c(
  "siteid", "dist_up_km", "ord_stra",
  "dis_m3_pyr", "ele_mt_cav", "slp_dg_cav",
   "hft_ix_c93", "hft_ix_c09"
)

riveratlas_tmp <- riveratlas_site %>%
  select(all_of(col_pca_riv)) %>%
  st_drop_geometry() %>%
  left_join(
    modelling_data %>%
      distinct(siteid, hft_ix_c9309_log2_ratio),
    by = "siteid") %>%
  select(-siteid) %>%
  mutate(hft_ix_c9309_ratio = hft_ix_c09 / hft_ix_c93) %>%
  rename_with(~str_replace_all(.,
        get_rev_vec_name_val(get_river_atlas_significant_var())
    )
  )

evt_summary <- riveratlas_tmp %>%
  pivot_longer(everything(),
    names_to = "Environmental variables",
    values_to = "values") %>%
  group_by(`Environmental variables`) %>%
  summarise(summ = list(enframe(summary_distribution(values)))) %>%
  unnest(summ) %>%
  pivot_wider(names_from = "name", values_from = "value") %>%
  mutate(
    median = round(median, 2),
    `Median (Q1, Q3)` = paste0(median, " (", round(`1st_quart`, 2), ",", round(`2nd_quart`, 2),
      ")"),
    `(Min, Max)` = paste0("(", round(min, 2), ",", round(max, 2), ")")
  ) %>%
  select(`Environmental variables`, N = n, `Median (Q1, Q3)`, `(Min, Max)`) %>%
  arrange(`Environmental variables`)

evt_summary %>%
  as_kable_extra(format = "latex")
```

```{r}
cap_pca_riv <- "Rotated PCA over the physical and hydrological structure of
streams. The first axis is related the average discharge, Strahler order and the
distance from source. This first axis was used in the statistical modelling as a
composite variable to summarise stream gradient from upstream to upstream."
```

```{r pca-riv-str, fig.cap=cap_pca_riv}
p_pca_riv_str <- plot_pca_clust(
  .data = pca_riv_str$rotated,
  site_cl = NULL,
  xaxis = "RC1", yaxis = "RC2",
  ctb_thld = .4, label_size = 2,
  size_abscisse_segment = .25,
  size_arrows_segment = .25,
  force = 10, force_pull = 1,
  var_scaling_factor = NULL,
  seed = NA,
  replace_var = get_rev_vec_name_val(get_river_atlas_significant_var()),
  add_variable = TRUE,
  add_point = FALSE,
  add_ellipse = FALSE
)
p_pca_riv_str
```

```{r}
lc_hft <- riveratlas_site %>%
  st_drop_geometry() %>%
  select(all_of(c(
        get_land_class_var()[str_detect(names(get_land_class_var()), "catchment")],
        get_river_atlas_significant_var()[
          str_detect(get_river_atlas_significant_var(), "hft_")
          ]
        )))
```

```{r}
cor_lc_hft <- cor(lc_hft,
  method = "spearman"
)
#corrplot::corrplot(cor_clust, type = "upper", diag = FALSE, tl.srt = 15)
p_cor_lc_hft <- ggcorrplot::ggcorrplot(
  cor_lc_hft, hc.order = TRUE, type = "lower",
  lab = TRUE, tl.srt = 15, legend.title = "Spearman Corr") +
theme(legend.position = c(.45, .95), legend.direction = "horizontal") +
guides(colour = guide_legend(title.position = "top"))
```

```{r cor-lc-hft, fig.cap="Spearman correlation between land uses and human footprint indexes"}
p_cor_lc_hft
```

# Statistical analysis

## Variable

```{r}
variable_stat <- tibble(
  category = c(rep("Response variable", 10), rep("Predictor variable", 7)),
  variables = c(clust_var, "perc_exo_sp", "perc_exo_abun", "log1_year_nb",
    "riv_str_rc1", "hft_ix_c93", "hft_ix_c9309_log2_ratio", "unitabundance",
    "main_bas", "siteid"),
  names = c(get_model_term_replacement_paper_figure(),
    get_var_replacement_vulgarisation())[variables],
  var_type = c("Factor", "Continuous", rep("Proportion", 8), "Continuous",
    "Continuous", "Continuous", "Proportion", "Factor", "Character",
    "Character"),
  unit = c("Count (reference), Number of individuals by 100 sq meters,
    Catch per Unit Effort (CPUE), Leslie index", rep("", 8), "Number of
  year", "", "", "", "Factor with 4 levels", "", "", ""),

  signification = c(
    "Total number of individuals",
    "Coverage corrected richness",
    "Temporal dissimilarity
    (presence/absence)",
    "Temporal dissimilarity
    (relative abundance, Simpson index)",
    "Temporal species
    colonisation",
    "Temporal species
    exptirpation", "Temporal species
    replacement", "Temporal
    nestedness",
    "Proportion of
    non-native species richness",
    "Proportion of
    non-native species abundance", "Time",
    "PCA axis related to average flow,
    distance from source, Strahler order, stream size",
    "Human footprint (1993)",
    "Human footprint ratio
    between 2009 and 1993",
    "Unit of abundance", "Hydrographic basin", "Sampled site"
    ),
    transformation = c("Log", "Log", rep("", 8), "Log + 1", "", "", "Log base
      2", "", "", "")
)

variable_stat %>%
  select(-variables, `short name` = names, type = var_type) %>%
  rename_with(str_to_sentence) %>%
  kbl(booktabs = T,
    label = "tab-var-mod",
    caption = "Description of the variables included in the model.") %>%
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle")
  kable_styling(latex_options = "striped")
```


## Collinearity

```{r}
vif_caption <- "Variance inflation factors on the variables included in the
model. SE factor: inflation of the standard error of slope coefficients
predicted by the multicollinearity of the variables"
```


```{r vif}
dummy_vif <- lm(log_species_nb ~
   log1_year_nb + riv_str_rc1 +
     hft_ix_c9309_log2_ratio +
    hft_ix_c93, modelling_data)
ck_col_data <- check_collinearity(dummy_vif) %>%
  as_tibble() %>%
  mutate(Term = str_replace(get_model_term_replacement()[Term], "\n", " ")) %>%
  rename(`SE factor` = "SE_factor", "Predictive variables" = "Term")
```
```{r}
ck_col_data %>%
  kbl(booktabs = T,
    label = "tab-vif",
    caption = vif_caption
    ) %>%
  kable_styling(latex_options = "striped")
  kable()
```


The presence of multicollinearity among explicative variables was checked by
measuring the Variance Inflation Factor (VIF) on a model formulation containing only
the main effect, i.e. not the interactions. We did so because main variables ($X_1$,
$X_2$) and the interactions are collinear by construction, interactions being the
products of main variables (i.e. $X_1X_2$). The VIF values were all close to 1
(Table \@ref(tab:vif)), indicating absence of multicollinearity.

## Selection of time modelling

```{r}
tar_load(comp_log_year_no_evt_re)
cap_aic_table <- "AIC of the models fitting either temporal trends with the
number of year since the start of sampling or the log of the former variable
added to one."
```

```{r aic, fig.cap=cap_aic_table}
aic_table_no_evt_re <- comp_log_year_no_evt_re %>%
  select(year_var, response, fit_summary) %>%
  unnest(fit_summary) %>%
  select(year_var, response, AIC) %>%
  mutate(AIC = round(AIC)) %>%
  pivot_wider(names_from = "year_var", values_from = "AIC") %>%
  mutate(
    which_min = map2_chr(year_nb, log1_year_nb, ~c("year_nb", "log1_year_nb")[which.min(c(.x, .y))]),
    diff_ratio_log_raw = (log1_year_nb - year_nb) / year_nb
  )
```

```{r}
aic_table_no_evt_re %>%
  mutate(
    response = get_var_replacement()[response],
    which_min = get_model_term_replacement()[which_min]) %>%
  rename(
    `Response variables` = response, `Min AIC` = which_min,
    `Proportion difference AIC` = diff_ratio_log_raw,
    `Year nb` = year_nb, 
    `Log (Year nb + 1)` = log1_year_nb
  ) %>%
  kbl(booktabs = T,
    label = "tab-aic",
    caption = cap_aic_table
    ) %>%
  kable_styling(latex_options = "striped")
```


## Model validity

```{r}
tar_load(obs_fit)
tar_load(gaussian_inla_exo_std)
cpo <- gaussian_inla_exo_std %>%
  mutate(
    cpo_pit = map(mod, ~inla.cpo(.x)),
    cpo = map(cpo_pit, ~.x$cpo$cpo),
    pit = map(cpo_pit, ~.x$cpo$pit)
    ) %>%
  select(-mod)
```

```{r}
obs_fit_cap <- "Fitted values by the models vs observed values. The black line
displays the bissection, i.e. the intercept and slope are respectively 0 and 1."
```


```{r fit-obs, fig.cap = obs_fit_cap}
obs_fit %>%
  filter(response %in% c(clust_var, "perc_exo_sp", "perc_exo_abun")) %>%
  mutate(response = get_var_replacement_vulgarisation()[response]) %>%
  ggplot(aes(x = obs, y = fit)) +
  geom_point(alpha = .3, color = "gray70") +
  geom_abline(slope = 1, intercept = 0, size = 1) +
  facet_wrap(vars(response), scales = "free") +
  labs(x = "Observed values", "Fitted values")
```

## Clustering of temporal trends

```{r}
tar_load(c(clust_curv_site, clust_curv_site_fac_50))
tar_load(c(k6_fac_1, sup_clust_size_fac1))
```

```{r}
cap_clust_curv <- "Objective function quantifying the goodness of k-means
clustering according the percentage of removed data (most outliers data in the
  multidimentional space) removed. In the results presented in the main text, we
removed 5% of the data. Top and bottom panels displayed results for restricting
factors of 1 and 50 respectively. Restricted factor of 1 was used for the main
text"
```


```{r, fig.cap=cap_clust_curv}
library(tclust)
plot_grid(
  plot(clust_curv_site),
  plot(clust_curv_site_fac_50),
  nrow = 2
)
```

```{r}
k4_cap <- "Display of the sites in the multidimentional space, colored by their cluster belonging. Number of clusters = 4"
```


```{r k4, fig.cap = k4_cap}
k4_fac_1_cl <- tclust(
  x = site_no_drivers_inla_tot[, clust_var] %>%
  mutate(across(everything(), ~scale(., center = FALSE)[,1])),
  iter.max = 150,
  k = 4,
  alpha = 0.05,
  restr.fact = 1,
  warnings = 2
)
plot(k4_fac_1_cl)
```

```{r}
bp_cl_dist_cap <- "Distribution of scaled temporal trends by variable and
cluster, for four clusters. With four cluster instead of six in the main text,
the clustering groups temporal trends of species richness and of total
abundance and the cluster characterising low change (Fig. 2, main text) is not
present."
```


```{r bp-k4, fig.cap=bp_cl_dist_cap}
site_k4_fac_1_cl <- get_cluster_df(
     tclust_obj = k4_fac_1_cl,
     site_env = site_env,
     assign_threshold = .5,
     clean_method = "rm"
     )
target_bp_cl_dist(cl_obj = site_k4_fac_1_cl) +
  theme(axis.text.x = element_text(angle = 20, vjust = 0))
```


# Supplementary results
 
## Figure with all the biodiversity facets

```{r, fig.show = "hide"}
tu_full <- rbind(gaussian_inla_std_effects,
  gaussian_inla_exo_std_effects) %>%
  filter(!str_detect(term, "unitabundance|Intercept")) %>%
  mutate(
    term = str_replace_all(term, get_model_term_replacement()),
    response = get_var_replacement()[response]
    )

p1_f <- tu_full %>%
  filter(facet == "main") %>%
  arrange(desc(width_bar)) %>%
  plot_inla_fixed_effect() +
  scale_x_break(c(.21, .5))

p2_f <- tu_full %>%
  filter(facet == "interaction") %>%
  arrange(desc(width_bar)) %>%
  plot_inla_fixed_effect()

p3_f <- tu_full %>%
  filter(facet == "dbl_interaction") %>%
  arrange(desc(width_bar)) %>%
  plot_inla_fixed_effect(.,
    legend_present = TRUE,
    xaxis_title = TRUE
    ) +
  labs(x = "Standardized coefficients")

leg_dis_f <- get_legend(p3_f +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow = 4))
)

figS1 <- plot_grid(
  print(p1_f),
  print(p2_f),
  p3_f + theme(legend.position = "none"),
  leg_dis_f,
  rel_heights = c(1, 1, 1, .2),
  ncol = 1)
```

```{r std-tot, fig.height = 12}
figS1
```

## Unstandardized coefficients

## Predictions of the model

```{r}
tar_load(c(pred_gaussian_inla, pred_gaussian_inla_exo))

pred_gaussian_inla_tot <- rbind(
  pred_gaussian_inla %>%
    filter(response %in% clust_var),
  pred_gaussian_inla_exo
  )
```

```{r, }
for (i in seq_along(pred_gaussian_inla_tot$response)) {
  p_list <- map(pred_gaussian_inla_tot$pred_plot[[i]],
    ~ .x + theme(legend.position = "bottom")
  )
  plot_grid(
    plotlist = p_list,
    ncol = 3)
}
```
